### 2011-11-21 wp1description/src/w1p.description.tex
Sections 1 and 2

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\title{Work Package 1 detailed description}
\date{\today}
\author{Peter Bloem}

\begin{document}

\maketitle

\begin{abstract}
A concise, non-formal description of the aims of Work Package 1 of Project 23
(Data2Semantics) of the COMMIT project. This documents provides the basic
intuitions required to understand the questions which this projects hopes to
answer.
\end{abstract}

Consider what we might mean by a `dataset'. A fMRI image of a subject's brain,
taken while the subject performs a specific task, a large body of annotated
language, TIFF images of the Taj Mahal, a collection of chess positions which
end in a draw. To be able to talk about data in the most general sense, we
suggest that any dataset can be translated to a single string of ones and zeroes
(as it is when stored on a computer system). We then build our basic concepts on
this description of data, so that we can be sure that they apply to everything
that we call data.

Any process, from simple operations like sorting a list of numbers, to learning
natural languages or discovering scientific knowledge, can be thought of as a
procedure on such a string of ones and zeroes. Any procedure that can be
described unambiguously and performed independent of any human interpretation or
ill-understood processes is called \emph{effective}. The notions of data encoded
as ones and zeroes, and effective procedures on such data have been thoroughly
studied over the last century.

This project will ask the following questions of such data:
\begin{itemize}
  \item Can we say, a priori, that some datasets are more interesting than
  others
  \item Can we describe a context in the broadest possible terms and define
  the interestingness or utility of a dataset withing such a context?
  \item What happens to the scientific process as we move into an era of
  increasingly large datasets? Will all scientific problems be solved simply by
  fast enough computer programs and large enough datasets, or will we find new
  limits that will defeat supercomputers and petabyte-sized datasets?
\end{itemize}

\section{Question 1: Complexity and Interestingness}

Consider the  following two datasets:
\indent101010101010101010101010101010101010101010101010101010101010101010101010101010\ldots
\indent101000110111111001110001011111010000001000101000110000111101010100101001100001\ldots
These examples represent extremes on a spectrum of interestingness. They are
sometimes called \emph{the clock} and \emph{the coinflip}. The first (the clock)
shows a high degree of regularity. It's perfectly predictable. The second shows
the highest possible degree of `chaos'. It's perfectly unpredictable.

There are many well defined notions of complexity. These measures tend to consider the coinflip maximally
complex and the clock maximally simple. While this certainly captures, for
instance, the effort required to describe the data, intuitively, it leaves
something to be desired.

Consider an old-fashioned black-and-white television. If we leave the
television off, the screen remains black. This is analogous to the clock above.
If we don't tune the receiver to a specific channel, we get a uniform static. By
most notions, he most complex signal the television is capable of displaying. It
is also commonly said that this signal contains the highest possible level of
information. That is were our intuition starts to confict with the theory.

Because while the static contains a lot of information in the sense that it is
difficult to describe exactly in simple terms, it is of no use to us. We are
much more interested in the television programs we see when we tune the dial to
a specific frequency. These are not entirely predictable and not entirely
unpredictable. We tend to be interested in the phenomena that lie somewhere
between the clock and the coinflip. Maximal information isn't enough, there must
be such a thing as `useful information'.

Intuitively, we guess that any datasets near the extremes will never be
interesting to anyone. How far can we push this? Can we narrow the band of
strings that are likely to be interesting? Can we suggest an optimum of a priori
interestingness somewhere between the clock and the coinflip? How far can we
take this without defining any kind of a context against which to measure
interestingness.

\section{Question 2: Conditional Complexity}

A priori interestingness can only go so far. A TIFF image of the Mona Lisa
may have a high a priori interestingness, but it's unlikely to to be of any use
to a theoretical physicist.

Determining interestingness on the basis of a context is easy to do in domain
specific terms. Given a well-formulated research question in molecular biology,
it should be simple enough to evaluate the utility of any dataset in answering
the research question, provided that the dataset is written in the same
`language' as the question. If we take the notion of `well-formed' questions to
its logical extreme, we end up with a computer program to be run on a dataset. A
query.

Can we generalize again? Ignore domain specific coding and programming and
simply consider the landscape of all datasets and all queries? The notions of
\epmh{data} and \emph{computer program} have been thoroughly generalized. The
concept of a question, or a query has (to our knowledge) not received the same
treatment.

\section{Question 3: Convergence}

\end{document}