### 2011-11-24 random/src/random_sets.tex
Stub

\documentclass{article}

\usepackage{charter}

\title{Investigating the random set}
\date{\today}
\author{Peter Bloem}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\begin{abstract}
\noindent Brief, formal definitions of various flavors of Kolmogorov Complexity, and an investigation of the set of random strings it defines.
\end{abstract}

\end{document}