### 2011-11-24 power laws/src/power_laws.tex
Stub

\documentclass{article}

\usepackage{charter}

\title{Proper power laws}
\date{\today}
\author{Peter Bloem}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\begin{abstract}
\noindent A thorough definition of the various types of power laws that exist, their realtions to one another and the general properties.
\end{abstract}

\end{document}