### 2011-11-24 power laws/src/power_laws.tex
Outline

\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}

\title{Power laws and other fat tails}
\date{\today}
\author{Peter Bloem}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\begin{abstract}
\noindent A thorough definition of the various types of power laws that exist, their realtions to one another and the general properties.
\end{abstract}

\section*{Power laws in general: functions with negative exponentials}
\subsection*{Pure power laws}

\[
f(x) = x^{-\alpha}
\]

\subsection*{Approximate power laws}

\[
f(x) = \phi(x)\;\cdot\;x^{-\alpha}
\]

\subsection*{Broken power law}

\subsection*{Exponential cutoff}

\section*{Power law distributions}

\subsection*{Attribute-based}

\subsubsection*{Pareto distribution}

\subsection*{Rank: probability as an attribute}

\subsubsection*{Zipf distribution}

\section*{Testing for power laws}

\section*{Fat tailed distributions}

\section*{Examples and applications}

This section provides some examples of power law and fat-tail phenomena and classifies them more exactly according to the definitions in this document.

\section*{Dimension}

\end{document}