### 2011-12-05 random/src/random_sets.tex
Added timesheet

\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem*{thm}{Definition}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{Kolmogorov Complexity}
\date{\today}
\author{Peter Bloem}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle

\begin{abstract}
\noindent This document provides a brief description of Kolmogorov Complexity and Marting L\o:f randomness. It may do some experimental tweking of notation and traditional definition, to highlight the important points in the development of Kolmogorov complexity that are often glossed over.
\end{abstract}

\section*{Notation and preliminaries}

$\cal{B}$ is the set of finite binary strings, including the empty string $\epsilon$. Members of $\cal B$ tend to be named $x$, $y$ and $z$.

A subset of $\cal B$ has the \textbf{prefix-property} if no member of the set is a prefix of another member. We fix one such subset $\cal P$ as a canonical example. Members of $\cal{P}$ tend to be named $\bar{x}$, $\bar{y}$, $\bar{z}$. \footnote{Note that this is different from the convention of having $\bar{x}$ be the prefix free encoding of $x$.}

If we fix some canonical ordering of $\cal B$, we can match each finite binary string to a natural number, thus ordering the binary string. We will sometimes consider the two interchangeable. Note that this also serves as an ordering for all subsets of $\cal A$. Let ${\cal A}_i$ be the $i$-th member of $\cal A$. For any binary string $x$, let ${\mathbb N}_x$ be the corresponding natural number by the agreed ordering.

Each subset of $\cal{B}$ encodes binary strings, we name this function after its set as follows: if ${\cal A} \subseteq {\cal B}$ then ${\cal A}({\cal B}_i) = {\cal A}_i$. Thus, ${\cal P}(x)$ is the prefix encoding of $x$.

If $x$ and $y$ are binary strings, then $xy$ represents the concatenation of the two.

\section*{The universal Turing machine}

\begin{dfn}
A \textbf{pairing function} $\langle\,\,\cdot,\,\cdot\,\rangle_f$ is any function $f: \mathbb{N} \times \mathbb{N} \rightarrow \mathbb{N}$ which is \emph{primitive recursive} and \emph{invertible}.
\end{dfn}

\begin{exm} The G\"{o}del pairing function is defined as $\langle a, b \rangle_g = 2^a3^b$. Since $2$ and $3$ are prime, each $a$ and $b$ will result in a product of a different sequence of primes, and thus a unique integer.
\end{exm}

\begin{exm} The \em{prefix pairing function} $\langle a, b \rangle_p = {\cal P}(a)b$ for any set ${\cal P}$ with the prefix property. Since a is prefix encoded, there is only one pair of inputs to each output. [Is this actually primitive recursive?]
\end{exm}

\begin{thm}
The set of all pairing functions is recursively enumerable.
\end{thm}

\begin{dfn}
A Turing machine $U$ is called universal if there is some pairing function $f$ and enumeration ${T_i}$ of Turing machines, such that if $x = \langle i, p\rangle$ then $U(x) = T_i(p)$. That is, a universal Turing machine can simulate any other turing machine running any other program.
\end{dfn}

If we think of Turing machines as computing a sequence of elements from ${\mathbb N} \cup \{\infty\}$, then a universal Turing machine weaves all these sequences together in a way that is easily decoded.

\end{document}