### 2014-10-17 Facticity/Facticity/facticity.tex

\documentclass{style/llncs}

\usepackage{eulervm}
\usepackage{charter}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{hyperref}
\usepackage{footmisc}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\title{Ideal model selection is impossible}

\author{Peter Bloem and Steven de Rooij}

% Steven's comments
\newcommand{\sdr}[1]{\textcolor{blue}{\small #1\textsuperscript{[sdr]} }}
% Peter's comments 
\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}
% Pieter's comments 
\newcommand{\pa}[1]{\textcolor{Red}{\small #1 \textsuperscript{[pa]} }}

\newcommand{\p}{\mbox{\,.}}
\newcommand{\fl}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\g}[1]{\color{gray} #1 \color{black}}
\newcommand{\B}{{\mathbb B}}
\newcommand{\Rbb}{{\mathbb R}}
\newcommand{\N}{{\mathbb N}}
\newcommand{\C}{{\mathscr C}}
\newcommand{\Co}{{\overline{\mathscr C}}}
\newcommand{\R}{{\mathscr R}}
\newcommand{\m}{{\overline{m}}}
\newcommand{\ok}{{\overline{\kappa}}}
\newcommand{\s}[1]{{\overline{#1}}}
\newcommand{\mmid}{\;\middle|\;}
\newcommand{\hide}[1]{}

\begin{document}

\maketitle

\begin{definition}[Kolmogorov complexity]
\[
K^\psi(x) = \min\left\{|\bar\imath p| : \psi_i(p) = x\right\}
\]
\end{definition}

\begin{definition}[Complexity of a function]
\[
K^{\phi\psi}(f) = \min\left\{|\bar\imath p| : \psi_i(p) = j, \forall_x\phi_j(x) = f(x)\right\}
\]
Where $\psi$ and $\psi$ represent acceptable numberings, and $f$ is some partial computable function.
\end{definition}

\begin{definition}[Specific Kolmogorov complexity]
\[
K^{\phi\psi}(x) = \min\left\{K^{\phi\psi}(f) + |p|: f(p) = x\right\}
\]
\end{definition}
This quantity is the same as the Kolmogorov complexity up to a constant. It also provides the value which all minimal two-part descriptions in our framework `witness', ie. it is the length of the shortest description in our two-part coding framework.

\begin{definition}[Sophistication]
\[
F^{\phi\psi}(x) = \min\left\{K^{\phi\psi}(f) : \exists_p f(p) = x, K^{\phi\psi}(f)+|p| \leq K^{\phi\psi}(x)\right\}
\]
where $\phi$ and $\psi$ represent acceptable numberings.
\end{definition}

\begin{lemma}
There exist numberings $\phi$ and $\psi$ such that $F^{\phi\psi}(x) \leq c$ for all $x$ and some constant $c$ independent of $x$. \label{lemma:constant}
\end{lemma}
\begin{proof}
Let $\psi$ be some acceptable numbering. We then define $\phi$ as
\begin{align*}
\phi_0(\bar\imath\jmath p) &= \psi_{\psi_i(j)}(p) \\
\phi_{11r\bar\imath}(p) &= \psi_(p)
\end{align*}
with $r$ a random string. For any $\phi_i$ where $i$ does not fit either of these cases, we define $\phi_i(x) = \infty$.

Let $x$ be any binary string. Let $\psi_i$ be some function with minimal $i$ such that $\psi_i(p) = x$. This gives us two representations in $\phi$: $\phi_0(\bar\imath j p)$ and $\phi_{ri}(p)$. These have the following description lengths:

\begin{description}
\item[$\phi_0(\imath j p)$:] $K^{\phi\psi}(\phi_0) + |\overline{a}bp| = K^\psi(0) + K^\psi(i) + |p|$
\item[$\phi_{ri}(p)$:] $K^{\phi\psi}(\phi_{ri}) + |p| = K^{\psi}(ri) + |p|$
\end{description}

The difference between the two is $K^{\psi}(11ri) - K^\psi(0) - K^\psi(i)$. For some $r$ large enough this will always be larger than $0$, which means that the two part coding using $\phi_0$ will always provide a smaller description. 
This means that $F^{\psi\phi}(x)$ will be $K^{\phi\psi}(\phi_0)$ for any $x$.
\end{proof}

\begin{lemma}
There exist a $\phi$ and $\psi$ such that for all $x$ with $K^\psi(x) > |x|$ (with $c$ some constant), we have $F^{\phi\psi}(x) > |x| - O(1)$. \label{lemma:unbounded}
\end{lemma}
\begin{proof}
Let $\psi$ be some acceptable numbering. We then define $\phi$ as
\begin{align*}
\phi_0(\bar\imath\jmath p) &= \psi_{\psi_i(j)}(p) \\
\phi_{11ri}(p) &= \psi_i(p)
\end{align*}

with $r$ a random string. For any $\phi_i$ where $i$ does not fit either of these cases, we define $\phi_i(x) = \infty$.

Let $x$ be a string with $K^\psi(x) > |x|$ as in the statement of the lemma. Let $\psi_i(p)$ be a smallest description under $\psi$. We have two descriptions in $\phi$:

\begin{description}
\item[$\phi_{0\bar\imath p}(\epsilon)$:] $K^{\phi\psi}(\phi_{0\bar\imath p}) K^{\phi}(0\bar\imath p) \leq |0\bar\imath p| + c$
\item[$\phi_{11ri}(p)$] $K^{\phi\psi}(11ri) + |p| = $  
\end{description}

Again, if $r$ is large enough, the first description will be shorter. Since $K^\psi(x) > |x|$, so isthe length of the model: $|0\bar\imath p | > |x|$, which means that the sophistication $F^{\psi\phi}$ grows linearly with $|x|$.

\end{proof}
\begin{theorem}
There exist pairs of acceptable numbers $\phi$, $\psi$ and $\eta$, $\mu$ such that for any constant $c$ there exist $x$ for which $\left|F^{\psi\phi}(x) -F^{\phi\psi}(x)\right| > c$ 
\end{theorem}
\begin{proof}
This follows directly from Lemmas~\ref{lemma:constant} and \ref{lemma:unbounded}.
\end{proof}

\bibliographystyle{style/splncs}
\bibliography{facticity}

\end{document}