### 2014-10-22 Nickname problem/src/nickname.tex

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath, amsthm, amssymb}

\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\title{The Nickname Problem}
\date{\ldots}
% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\begin{document}

\maketitle
The \textbf{nickname problem} arises when we use an indexing of the computable functions as a mathematical representation of all possible models, and their complexity.

Consider, for instance, the string $\sigma$ representing the collected works of Shakespeare in binary form. We can define a simple enumeration of the recursive functions based on Turing Machines as usual, with the single change that a Turing Machine which outputs $\sigma$ (disregarding its input) is given index 20. While pathological, this index is perfectly computable, and $\sigma$ would be seen as a highly compressible string, consisting purely of model information.

We might also construct an enumeration which places a Turing machine at index 0 which produces the collected works of Shakespeare only for a high input with slightly smaller than any other description of $\sigma$. Under this enumeration, the string would be seen as reasonably compressible, but with only one bit of model information.

The nickname problem is inescapable. Kolmogorov Complexity itself suffers from a similar problem, where for a given string a pathological universal machine can be constructed which compresses that string to a single bit. Like with Kolmogorov complexity, we would like some assurance that this kind of pathological fiddling can only affect a limited number of string in a limited way. In other words, we can change the nature of facticity by switching to a different universal machine, but only to a limited extent.

\begin{dfn}
For a given index of the computable functions $\{\phi_i\}$ we define $i^*_\phi = \min\left\{\phi_j : \phi_j \simeq \phi_i\right\}$. That is, the minimal index of a function that is equivalent to $\phi_i$. The subscript may be dropped when it is clear from context.
\end{dfn}

\begin{dfn}
An index of the computable functions $\{\phi_i\}$ is called \textbf{compact} if for every 
$\phi_i(\bar{x}y)$ we have a $\phi_j$ such that $\phi_i(\bar{x}y) \simeq \phi_j(y)$ and  $|j| - |i| \leq |\bar{x}| + O(1)$
\end{dfn}

Note that this is a natural property of basic computational metaphors. We can always take part of the argument of one function and hardcode it into the program at the cost of its prefix-coded length, and some additional bits to add it to the argument before processing.

We can construct a non-compact enumeration by starting with a compact one, and inserting after function $i$ a series of $f(i)$ copies of $\phi_0$, where $f(i)$ is some fast growing function like $2^{i!}$.

\begin{lma}
Let $\{\phi_i\}$ and $\{\psi_j\}$ be two compact, computable enumerations of the computable functions. For a given computable function $\phi_i \simeq \psi_j$,  we have $\mbox{abs}(|i^*| - |j^*|) \leq \max(|\overline{\imath^*}|, |\overline{\jmath^*}|) + O(1)$. That is, the difference between the sizes of the minimal models is in $O(\log(\max(i, j)))$
\end{lma}
\begin{proof}
Let $\phi_u$ be a universal function for $\psi$, such that $\phi_u(\bar{\imath}x) \simeq \psi_i(x)$. Since $\psi$ is computable this function exists in $\phi$. We have $\psi_{j^*}(x) \simeq \phi_u(\overline{\jmath^*}x)$. By the compactness of $\phi$, we have some $h$ with $|h| \leq |u| + |\overline{\jmath^*}| + O(1)$ such that $\phi_{h}(x) \simeq \phi_u(\overline{\jmath^*}x) \simeq \psi_{j^*}(x)$.

Thus $|i^*| \leq |h| = |u| + |\overline{\jmath^*}| + O(1)$. Reversing $\phi$ and $\psi$ gives us $|j^*| \leq |h'| = |u'| + |\overline{\imath^*}| + O(1)$. Combining these, we get the desired result.
\end{proof}

This lemma tells us that so long as we restrict our indices to compact and effective, we can never gain more than a logarithmic difference between model sizes. In the limit as $i \rightarrow \infty$, this term represents a vanishing proportion.

\end{document}