### 2015-10-12 fractal paper/src/fractal.tex

\documentclass[10pt,a4paper,oneside]{article}

\usepackage{amsmath, amsthm, amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\theoremstyle{definition}
\newtheorem*{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem*{dfn}{Definition}
\newtheorem*{exm}{Example}

\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\p}{\mbox{\,.}}
\newcommand{\bD}{\boldsymbol\Lambda}
\newcommand{\bm}{\boldsymbol\mu}
\newcommand{\cN}{{\cal N}}
\newcommand{\bs}[1]{{\boldsymbol {#1} }}

\title{A variational algorithm for the Fractal Inverse Problem.}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\newcommand{\pb}[1]{\textcolor{OliveGreen}{\small #1 \textsuperscript{[pb]} }}

\begin{document}

\maketitle

\begin{abstract}
\noindent We present an algorithm for solving the \emph{inverse problem} for iterated function systems (IFS) and random iterated function systems (RIFS). The central idea is to cast the IFS as an generalization of a mixture of of Gaussians model, and to treat the particular sequence of components that leads to a particular point as a latent variable.
\end{abstract}
\section{Introduction}
\pb{Briefly: importance of fractals, inverse problem.}

\pb{Iterated Function Systems.}

\pb{Idea behind our approach. EM principle, using variational Bayes.}

\subsection{Preliminaries}
We will deal with datasets which are a set of points $\{x_1, \ldots, x_n\}$ with $x_i \in \mathbb{R}^D$.

Let $F_{t, R}(y) = t + Rx$ be an affine transformation represented by a vector $t$ and a matrix $R$. If $X ~ \cN(\mu, \Sigma)$, then $F_{t, R}(Y) ~ \cN(t + R\mu, R\SigmaR^T)$. Specifically this means that if we take the standard multivariate normal $\cN_0 = \cN(0, I)$ as  canonical source, we can represent every multivariate normal $\langle \mu, \Sigma \rangle$ by an affine transformation $\langle t, R \rangle$ with $t = \mu$, $\Sigma = RR^T$. It also tells us that sampling from $\cN_0$ and applying a series of affine transformations is equivalent to sampling from a particular MVN.

\paragraph{Variational Bayes} Let $x$ be the observed data of a problem, for which we have a model, and let $z$ be the latent variables of the model. If the model has any parameters, we place priors on them so that we can consider them latent variables, and absorb them into $z$ The variational Bayes approach attempts to approximate the distribution $p(z \mid x)$ using a \emph{variational distribution} $q(z)$. We assume that $z$ can be split into subsets $z_1, \ldots, z_m$, for which $q$ factorizes:
\[
q(z) = \prod_i q_i(z_i)
\]
A central result in variational Bayes is that the optimum for each factor is given by:
\begin{equation}
q^*_i(z_i) = \langle \ln p(x, z)\rangle_{/i} + \text {const} \p \label{line:var-bayes}
\end{equation}
Where $\langle \ln p(x, z) \rangle_{/i} = \sum \prod_{j = {1, \ldots, m}/i} q_j(z_j) ln p(x, z)$ with the sum over all values of all $z_j$ except $z_i$. That is, it represents the expectation with respect to all latent variable factors except $z_i$.

Usually, the values of these optimal factors have cyclical dependencies, which means that we cannot compute $q^*$ directly. We can, however, search for a local optimum by iteration: we start with some initial value, and use (\ref{line:var-bayes}) to compute new values for each factor based on those of the previous generation.

\section{The IFS model}

\begin{figure*}[t]
  \includegraphics[width=\textwidth]{../img/factor-graph.pdf}
  \caption{A graphical model, illustrating the components of the IFS model. The gray box is a plate, representing a repetition of the nodes inside the box for each datapoint. The black node represents the observed data.}
  \label{figure:ifs-diagram}
\end{figure*}

The complete IFS model is illustrated in Figure~\ref{figure:ifs-diagram}. If contains the following components:
\begin{description}
\item[the component weights $\pi$] This is a vector of $k$ nonnegative real values, with $\sum_i \pi_i = 1$. Each weight determines the probability of its associated component being chosen at each iteration of the IFS.
\item[the depth $d$] The number of times the IFS is iterated.
\item[the components $\bm = \{\mu_1, \ldots, \mu_k\}$, $\bD = \{\Lambda_1, \ldots, \Lambda_k\}$] The means and precision matrices for the $k$ components. 
\item[The post-transformation $\mu_\text{post}$, $\Lambda_\text{post}$] The parameters for a single post-transformation. To see why this is necessary, consider the Koch curve in Ficgure~\ref{figure{fractals}}. While a dataset sampled from this measure can be prefectly represented by an IFS, it is slighty off-center. If this were real dataset, it might be mean-centered instead, in which case no IFS would fit. Unlike the mixture of Gaussians model, an IFS cannot be translated by a straightforward change of the parameters,a nd as the Koch curve shows, simple solutions like mean-centering the data do not always solve the problem. For this reason, we add a single post-transformation into the model, to allow for any simple, affine transformations to the final limit measure. This factor has the additional advantage that at $d = 0$, the model collapses to a single MVN distribution.
\item[The transformation sequence $\sigma_i$] For each data point $x_i$, we have a sequence of length $d$ representing the sequence of component-transformations we need to apply to a point sampled from $\cN_\text{pre}$ to get the final point sampled from the limit measure of the IFS. 
\item[The data $x_i$] This is the point after the post-transformation is applied: ie. the observed data.
\end{description}

The sampling procedure modeled

This gives us the joint distribution $p(\bs{x}, \bs{\sigma}, \mu_\text{pre}, \Lambda_\text{pre}, \bm, \bD, \mu_\text{post}, \Lambda_\text{post}, \pi, d)$ which factorizes as follows:
\begin{align}
p(&\bs{x}, \bs{x'}, \bs{\sigma}, \mu_\text{pre}, \Lambda_\text{pre}, \bm, \bD, \mu_\text{post}, \Lambda_\text{post}, \pi, d) \notag\\ 
=\;& p(\mathbf{x} \mid \mathbf{x'}, \mu_\text{post}, \Lambda_\text{post}) \; p(\mu_\text{post}, \Lambda_\text{post}) \notag\\
& p(\mathbf{x'} \mid \bs{\sigma}, \bm, \bD, \mu_\text{pre}, \Lambda_\text{pre}) \notag\\
& p(\mathbf{\sigma} \mid \pi, d) \; p(\pi) \; p(d) \notag\\
& \prod_i p(\mu_i, \Lambda_i) \p \label{line:factorization}\\
\end{align}

We use the following priors:
\begin{align*}
p(\pi) &= \text{Dir}(\bs{\alpha}_0) \\
p(d) &= {\cal U}(0, d_\text{max}) \\
p(\mu, \Lambda) &= p(\mu \mid \Lambda) p(\Lambda) = \cN(\mu \mid \bs{m}_0, (\beta_0\Lambda)^{-1}) {\cal W}(\Lambda \mid \bs{W}_0, \nu_0) \p
\end{align*}

Where the last line, a Gaussian-Wishart prior, applies to each pair of $\mu$ and $\Lambda$.

\subsection{The Inference algorithm for IFS models}
We assume the following factorization for our variational distribution to approximate  $p(\bs{z}|\bs{x})$:
\begin{align*}
q(& \mu_\text{pre}, \Lambda_\text{pre}, \bm, \bD, \bs{\sigma}, \pi, d) \\ 
=\;&  q_{\mu_\text{pre}, \Lambda_\text{pre}}(\mu_\text{pre}, \Lambda_\text{pre}) \; q_{\bm, \bD}(\bm, \bD)\; q_{\bs{\sigma}}(\bs{\sigma}) \; q_{\pi, d}(\pi, d) \p \\
\end{align*}
For the sake of clarity, we will omit the subscripts from the factors in the following.

\subsection{Update rules}

\paragraph{Responsibilities}

The derivation of the update rule for the responsibilities, ie. $q_{bs{\sigma}}^{n}(\bs{\sigma})$ takes the same form as the case for the mixture of Gaussians model \cite{}, since our model is essentially a mixture of $k^d$ Gaussians. Note that we can visualize $q(\bs{sigma})$ as a large matrix with $n$ rows and $2^d$ columns.

We have:
\begin{align*}
q^n(\bs{\sigma}) &= \langle \ln p(x, \bm, \bD, \mu', \Lambda', \bs{\sigma}, \pi, d) \rangle_{\bm, \bD, \mu', \Lambda', \pi, d} + const \p
\end{align*}

We factorize $\ln p(\ldots)$ according to (\ref{line:factorization}), and absorb any terms that don't depend on $\bs{\sigma}$ into the constant:   
\begin{align*}
q^n(\bs{\sigma}) &= \langle \ln p(x \mid \bm, \bD, \mu', \Lambda', \bs{\sigma}) \rangle_{\bm, \bD, \mu', \Lambda'} +  \langle \ln p(\bs{\sigma} \mid \pi, d) \rangle_{\pi, d} + const \p\\
& = 
\end{align*}

\paragraph{Components}

\paragraph{Component weights}

\paragraph{Pre-transformation}

\paragraph{Depth}

\subsection{Results}

\section{The RIFS model}
\paragraph{Random Iterated Function Systems}

\subsection{The Inference algorithm for RIFS models}
\subsection{Results}

\section{Discussion}

\pb{Recap.}

\pb{}

\bibliographystyle{siam}
\bibliography{fractal}

\end{document}