### 2018-09-05 kgmotifs/supplement.tex

\documentclass[letterpaper]{article} %DO NOT CHANGE THIS

\usepackage{pdflscape}
\usepackage{times}     % Required
\usepackage{helvet}    % Required
\usepackage{courier}   % Required
\usepackage{url}       % Required
\usepackage{graphicx}  % Required
\frenchspacing         % Required
\setlength{\pdfpagewidth}{8.5in}  % Required
\setlength{\pdfpageheight}{11in}  % Required
\usepackage{longtable}
\usepackage{makecell}

% PDF Info Is Required:
\pdfinfo{
/Title (Supplement: Detecting Motifs in Knowledge Graphs using Compression)
/Author (...)}

\setcounter{secnumdepth}{0}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{float}

\usepackage{easylist}
\usepackage{soul}

\newcommand{\N}{{\mathbb N}}
\newcommand{\R}{{\mathbb R}}
\newcommand{\Z}{{\mathbb Z}}

\newcommand{\X}{{\cal X}}
\newcommand{\B}{{\mathbb B}}
\newcommand{\G}{{\cal G}}
\newcommand{\cS}{{\cal S}}

\newcommand{\I}{{\cal I}}
\newcommand{\tab}{\hspace*{5mm}}
\newcommand{\from}{\leftarrow}

\renewcommand\cellalign{tl}

\floatstyle{ruled}
\newfloat{pseudo}{h}{lop}
\floatname{pseudo}{Algorithm}

\title{Supplement: Detecting Motifs in Knowledge Graphs using Compression}

\begin{document}
\maketitle 
\section{The canonical isomorphism algorithm for patterns}

During search, we want to re-order the nodes and links in an encountered pattern to some canonical ordering. This way, when we later return an isomorphic pattern, we recognize it, and will not need to compute its logfactor again. 
The \emph{nauty} algorithm is the standard solution for findings such \emph{canonical isomorphs}. As we are sampling partially labeled, directed graphs, we have a lot of information that can speed up canonicalization. Since this step is usually the bottleneck in sampling, it's important to perform it as efficiently as possible. Below, we provide a very succinct description of how nauty works, when applied to our patterns.

Let $V'_M$ and $R'_M$ represent the labels of the variable nodes and links in $M$. Pattern $M$ is an isomorphism of pattern $M'$ if there exist bijections $\sigma_v : V'_M \times V'M$ and $\sigma_r: \R'_M \times \R'_M$ such that applying these to the variable elements of $M$ yields $M'$ (under graph equivalence). The set of all patterns isomorphic to $G$ is the isomorphism class $[G]$. A function $C(P)$ is a canonicalization function if it produces for all any pattern $M$ in $[M]$ the same, canonical member of $M$.

Note that a very simple canonicalization algorithm would be to pass the unlabeled graph to the Nauty algorithm, and use its canonical ordering on the labeled graph. However, the labels can make the canonicalization procedure significantly faster.

The key principle behind the Nauty algorithm is the use of a \emph{colouring}: a partition of the nodes of a graph. A \emph{safe colouring} is defined as a partition $\pi$ of the nodes of the graph such that any permutation that maps nodes from different partitions to one another is guaranteed \emph{not} to be an automorphism. No guarantees are made about nodes in the same cell. For instance, the partition that puts all nodes in the same cell is always safe. If two nodes have different degrees, we can put them in different cells and be sure that we have a safe colouring. Given one safe colouring $\pi$, if node $a$ from partition $i$ has a neighbor in partition $j$ and node $b$ from partition $j$ has no such node, we can \emph{refine} the partition $\pi$ by separating $i$ into a new partition (one contain $a$ and one containing $b$).

The nauty algorithm operates on a graph with a given colouring (with the unit colouring used by default). It first refines the colouring as much as possible, using degree information, and propagating the distinguishing features through the graph. Once the most refined colouring has been achieved, it searches through all autorphisms of the graph to produce a canonical example (using various optimizations to prune the search tree).

In our case, we use the colouring to indicate what our labeling tells us about the colouring of the nodes.
Note the following:
\begin{easylist}
§ Labeled nodes in the pattern can be placed in a discrete cell of the colouring.
§ Variable nodes should be placed into a single cell in the partition.
§ We define a partition over nodes \emph{and} relation labels. 
\end{easylist}

\begin{landscape}

\section*{Motifs found for the three datasets}

\subsection{Dogfood, top 100 by log factor}

\begin{longtable}{ r r p{18cm} }
 log factor & frequency & \\
\hline\endhead
 28232.1 & 5807 & \makecell{\texttt{?n1 swc:heldBy ?n2.} } \\ 
 5125.5 & 2538 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 3997.0 & 2180 & \makecell{\texttt{?n1 swc:isPartOf ?n2.} } \\ 
 1755.1 & 566 & \parbox{2cm}{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_tzid ?n2.} } \\ 
 378.1 & 16651 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 rdf:\_1 ?n3.} } \\ 
 201.2 & 539 & \makecell{\texttt{?n1 swc:hasRole ?n2.} \\\texttt{?n2 rdf:type swc:Presenter.} } \\ 
 -2723.1 & 466 & \makecell{\texttt{?n1 rdf:type swc:Presenter.} } \\ 
 -2994.2 & 2031 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -3191.3 & 30 & \makecell{dsw:conference/eswc/2006/sessions/poster-session swc:hasRole ?n1. \\\texttt{?n1 rdf:type swc:Presenter.} } \\ 
 -3227.7 & 154 & \makecell{\texttt{dsw:conference/iswc/2008/research/pcmember swc:heldBy ?n1.} } \\ 
 -3271.9 & 156 & \makecell{\texttt{dsw:conference/iswc/2008/research/pcmember ?p2 ?n1.} } \\ 
 -3415.6 & 49 & \texttt{dsw:conference/dc/2010/programme-committee-member swc:heldBy ?n1.} \\ 
 -3514.0 & 138 & \makecell{\texttt{?n1 ?p2 dsw:person/eero-hyvoenen.} } \\ 
 -3596.0 & 29 & \makecell{\texttt{dsw:conference/iswc/2008/in\_use/pcmember swc:heldBy ?n1.} } \\ 
 -3605.8 & 33 & \makecell{\texttt{?n1 ?p2 dsw:person/heiko-paulheim.} } \\ 
 -3608.1 & 31 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member swc:heldBy ?n1.} } \\ 
 -3608.5 & 96 & \makecell{\texttt{\_:genid8 ?p2 ?n1.} } \\ 
 -3619.3 & 81 & \makecell{\texttt{\_:genid14 ?p2 ?n1.} } \\ 
 -3637.2 & 55 & \makecell{\texttt{\_:genid37 ?p2 ?n1.} } \\ 
 -3638.1 & 60 & \makecell{\texttt{\_:genid17 ?p2 ?n1.} } \\ 
 -3642.2 & 36 & \makecell{\texttt{dsw:conference/iswc/2010/dc/pcmember swc:heldBy ?n1.} } \\ 
 -3643.8 & 47 & \makecell{\texttt{?n1 ?p2 dsw:person/kim-viljanen.} } \\ 
 -3644.5 & 46 & \makecell{\texttt{?n1 ?p2 dsw:person/bernhard-schandl.} } \\ 
 -3656.0 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/paper/services/3/authorlist rdf:\_1 dsw:person/heiko-paulheim.} } \\ 
 -3658.5 & 1 & \makecell{\texttt{\_:genid12 rdf:\_10 dsw:person/michael-pool.} } \\ 
 -3660.9 & 1 & \makecell{\texttt{\_:genid14 rdf:\_1 dsw:person/siegfried-handschuh.} } \\ 
 -3660.9 & 19 & \makecell{\texttt{?n1 ?p2 dsw:person/xi-bai.} } \\ 
 -3662.3 & 1 & \makecell{dsw:person/doroteo-t-toledano foaf:name "Doroteo T. Toledano". } \\ 
 -3662.8 & 34 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member ?p2 ?n1.} } \\ 
 -3662.9 & 1 & \makecell{\texttt{dsw:person/doroteo-t-toledano rdfs:label "Doroteo T. Toledano".} } \\ 
 -3663.6 & 1 & \makecell{\texttt{dsw:conference/www/2012/demo/demo-chair swc:heldBy dsw:person/bebo-white.} } \\ 
 -3664.9 & 1 & \makecell{\texttt{\_:genid17 rdf:\_1 dsw:person/john-barton.} } \\ 
 -3665.5 & 1 & \makecell{dsw:organization/czech-technical-university-in-prague foaf:member dsw:person/petr-pollak. } \\ 
 -3665.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/431/authorlist rdf:\_1 dsw:person/alexandre-passant.} } \\ 
 -3666.2 & 1 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 dc:creator dsw:person/violeta-damjanovic.} } \\ 
 -3666.4 & 1 & \makecell{\texttt{\_:genid37 rdf:\_2 dsw:person/stephen-larson.} } \\ 
 -3666.7 & 1 & \makecell{\texttt{\_:genid117 rdf:\_1 dsw:person/tomas-vitvar.} } \\ 
 -3666.8 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember rdf:type swc:OrganisingCommitteeMember.} } \\ 
 -3666.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/468/authorlist rdf:\_1 dsw:person/jerome-euzenat.} } \\ 
 -3666.8 & 1 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 dc:creator dsw:person/raluca-zaharia.} } \\ 
 -3666.8 & 1 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 dc:creator dsw:person/laurentiu-vasiliu.} } \\ 
 -3667.2 & 1 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/43 foaf:maker dsw:person/malte-kiesel.} } \\ 
 -3667.2 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember swc:heldBy dsw:person/achille-fokoue.} } \\ 
 -3667.6 & 3 & \makecell{\texttt{dsw:conference/lrec/2008/papers/421 ?p1 dsw:person/petr-pollak.} } \\ 
 -3667.6 & 1 & \makecell{\texttt{dsw:conference/lrec/2008/papers/421 dc:creator dsw:person/petr-pollak.} } \\ 
 -3667.6 & 1 & \makecell{\texttt{dsw:workshop/iwod/2008/main/chair/organisingcommitteemember rdf:type swc:OrganisingCommitteeMember.} } \\ 
 -3667.7 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member swc:heldBy dsw:person/tomas-vitvar.} } \\ 
 -3667.8 & 1 & \makecell{\texttt{\_:genid39 rdf:\_1 dsw:person/tomas-vitvar.} } \\ 
 -3667.9 & 3 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 ?p1 dsw:person/raluca-zaharia.} } \\ 
 -3667.9 & 3 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 ?p1 dsw:person/laurentiu-vasiliu.} } \\ 
 -3668.0 & 1 & \makecell{\texttt{dsw:workshop/swsc/2008/main/pcmember swc:heldBy dsw:person/tomas-vitvar.} } \\ 
 -3668.1 & 59 & \makecell{\texttt{?n1 ?p2 dsw:person/achille-fokoue.} } \\ 
 -3668.3 & 1 & \makecell{\texttt{\_:genid38 rdf:\_1 dsw:person/nicola-fanizzi.} } \\ 
 -3668.3 & 1 & \makecell{\texttt{dsw:conference/lrec/2008/papers/301 swrs:author dsw:person/hiroyuki-shinnou.} } \\ 
 -3669.0 & 1 & \makecell{\texttt{\_:genid37 rdf:\_2 dsw:person/hanmin-jung.} } \\ 
 -3670.3 & 1 & \makecell{\texttt{dsw:conference/iswc/2008/research/pcmember swc:heldBy dsw:person/chiara-ghidini.} } \\ 
 -3670.6 & 13 & \makecell{\texttt{?n1 rdf:type swc:OrganisingCommitteeMember.} } \\ 
 -3670.9 & 80 & \makecell{\texttt{\_:genid12 ?p2 ?n1.} } \\ 
 -3671.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/dc/pcmember swc:heldBy dsw:person/andreas-harth.} } \\ 
 -3671.2 & 18 & \makecell{\texttt{\_:genid17 rdf:\_1 ?n1.} } \\ 
 -3673.4 & 3 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/43 ?p1 dsw:person/malte-kiesel.} } \\ 
 -3673.7 & 13 & \makecell{\texttt{?n1 ?p2 dsw:person/hiroyuki-shinnou.} } \\ 
 -3674.2 & 3 & \makecell{\texttt{dsw:conference/lrec/2008/papers/301 ?p1 dsw:person/hiroyuki-shinnou.} } \\ 
 -3674.7 & 69 & \makecell{\texttt{?n1 ?p2 dsw:person/andreas-harth.} } \\ 
 -3675.6 & 20 & \makecell{\texttt{?n1 ?p2 dsw:person/jianfeng-du.} } \\ 
 -3676.1 & 1 & \makecell{\texttt{dsw:person/akira-mori rdf:type foaf:Person.} } \\ 
 -3677.1 & 1 & \makecell{\texttt{dsw:person/alberto-musetti rdf:type foaf:Person.} } \\ 
 -3677.9 & 1 & \makecell{\texttt{dsw:person/alejandro-jaimes rdf:type foaf:Person.} } \\ 
 -3678.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/paper/services/3/authorlist ?p1 dsw:person/heiko-paulheim.} } \\ 
 -3678.7 & 13 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n1.} } \\ 
 -3678.8 & 1 & \makecell{\texttt{dsw:person/georg-lausen rdf:type foaf:Person.} } \\ 
 -3679.2 & 2 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 ?p1 dsw:person/violeta-damjanovic.} } \\ 
 -3679.5 & 10 & \makecell{\texttt{\_:genid8 rdf:\_4 ?n1.} } \\ 
 -3679.7 & 10 & \makecell{\texttt{\_:genid59 rdf:\_2 ?n1.} } \\ 
 -3680.2 & 2 & \makecell{\texttt{dsw:person/doroteo-t-toledano ?p1 "Doroteo T. Toledano".} } \\ 
 -3680.3 & 13 & \makecell{\texttt{\_:genid39 rdf:\_1 ?n1.} } \\ 
 -3680.5 & 1 & \makecell{\texttt{\_:genid12 ?p1 dsw:person/michael-pool.} } \\ 
 -3680.7 & 3 & \makecell{\texttt{dsw:conference/iswc/2009/paper/research/347 ?p1 dsw:person/bernhard-schandl.} } \\ 
 -3682.2 & 33 & \makecell{\texttt{\_:genid38 ?p2 ?n1.} } \\ 
 -3682.9 & 1 & \makecell{\texttt{\_:genid14 ?p1 dsw:person/siegfried-handschuh.} } \\ 
 -3683.8 & 13 & \makecell{\texttt{\_:genid37 rdf:\_2 ?n1.} } \\ 
 -3685.6 & 1 & \makecell{\texttt{dsw:conference/www/2012/demo/demo-chair ?p1 dsw:person/bebo-white.} } \\ 
 -3685.7 & 1 & \makecell{\texttt{\_:genid150 ?p1 dsw:person/bettina-bauer-messmer.} } \\ 
 -3686.3 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/60/authorlist ?p1 dsw:person/xi-bai.} } \\ 
 -3686.8 & 169 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n2.} \\\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid38.} } \\ 
 -3687.0 & 43 & \makecell{\texttt{\_:genid42 ?p2 ?n1.} } \\ 
 -3687.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/social\_web/pcmember ?p1 dsw:person/eero-hyvoenen.} } \\ 
 -3687.2 & 3 & \makecell{\texttt{dsw:conference/eswc/2008/paper/54 ?p1 dsw:person/eero-hyvoenen.} } \\ 
 -3687.5 & 1 & \makecell{\texttt{dsw:organization/czech-technical-university-in-prague ?p1 dsw:person/petr-pollak.} } \\ 
 -3688.1 & 1 & \makecell{\texttt{dsw:conference/rr/2008/full/4/authors ?p1 dsw:person/jianfeng-du.} } \\ 
 -3688.4 & 1 & \makecell{\texttt{\_:genid37 ?p1 dsw:person/stephen-larson.} } \\ 
 -3688.7 & 1 & \makecell{\texttt{\_:genid117 ?p1 dsw:person/tomas-vitvar.} } \\ 
 -3688.7 & 2 & \makecell{\texttt{\_:genid8 ?p1 dsw:person/irene-celino.} } \\ 
 -3688.8 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember ?p1 swc:OrganisingCommitteeMember.} } \\ 
 -3688.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/468/authorlist ?p1 dsw:person/jerome-euzenat.} } \\ 
 -3689.0 & 13 & \makecell{\texttt{\_:genid37 rdf:\_1 ?n1.} } \\ 
 -3689.1 & 1 & \makecell{\texttt{\_:genid36 ?p1 dsw:person/benjamin-coe.} } \\ 
 -3689.2 & 2 & \makecell{\texttt{dsw:conference/eswc/2011/paper/in-use/52 ?p1 dsw:person/eero-hyvoenen.} } \\ 
 -3689.3 & 1 & \makecell{\texttt{dsw:conference/eswc/2006/demo-papers/FD01-hitzler ?p1 dsw:person/malte-kiesel.} } \\ 
 -3689.3 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember ?p1 dsw:person/achille-fokoue.} } \\ 
 -3689.7 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/poster-track-committee-member ?p1 dsw:person/tomas-vitvar.} } \\ 
 -3689.8 & 8 & \makecell{\texttt{\_:genid96 rdf:\_1 ?n1.} } \\ 
 -3690.1 & 1 & \makecell{\texttt{dsw:workshop/swsc/2008/main/pcmember ?p1 dsw:person/tomas-vitvar.} } \\ 
 -3690.3 & 1 & \makecell{\texttt{\_:genid38 ?p1 dsw:person/nicola-fanizzi.} } \\ 
 -3691.0 & 1 & \makecell{\texttt{\_:genid37 ?p1 dsw:person/hanmin-jung.} } \\ 
 -3692.3 & 3 & \makecell{\texttt{?n1 swc:heldBy dsw:person/bebo-white.} } \\ 
 -3692.9 & 47 & \makecell{\texttt{\_:genid36 ?p2 ?n1.} } \\ 
 -3693.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/dc/pcmember ?p1 dsw:person/andreas-harth.} } \\ 
 -3694.6 & 1 & \makecell{\texttt{?n1 rdf:\_10 dsw:person/michael-pool.} } \\ 
 -3694.6 & 1 & \makecell{\texttt{\_:genid12 rdf:\_10 ?n1.} } \\ 
 -3695.0 & 2 & \makecell{\texttt{?n1 swc:heldBy dsw:person/marco-neumann.} } \\ 
 -3695.4 & 13 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n1.} \\\texttt{dsw:conference/iswc/2008/paper/poster\_demo/18 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid38.} } \\ 
 -3695.9 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/hiroyuki-shinnou.} } \\ 
 -3696.4 & 39 & \makecell{\texttt{\_:genid39 ?p2 ?n1.} } \\ 
 -3696.7 & 6 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/45 foaf:maker ?n1.} } \\ 
 -3697.1 & 1 & \makecell{\texttt{dsw:conference/www/2012/dev/16/talk swc:hasRole ?n1.} \\\texttt{?n1 rdf:type swc:Presenter.} } \\ 
 -3697.1 & 2 & \makecell{\texttt{dsw:person/doroteo-t-toledano foaf:made ?n1.} } \\ 
 -3697.6 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/bruno-alves.} } \\ 
 -3698.0 & 1 & \makecell{\texttt{dsw:conference/dc/2010/tutorial/2/tutorial-presenter rdf:type swc:Presenter.} \\\texttt{?n1 swc:hasRole dsw:conference/dc/2010/tutorial/2/tutorial-presenter.} } \\ 
 -3698.1 & 1 & \makecell{\texttt{dsw:person/akira-mori ?p1 foaf:Person.} } \\ 
 -3698.5 & 1 & \makecell{\texttt{?n1 foaf:name "Doroteo T. Toledano".} } \\ 
 -3698.8 & 1 & \makecell{\texttt{dsw:person/michalis-foukarakis ?p1 foaf:Person.} } \\ 
 -3699.0 & 1 & \makecell{\texttt{?n1 rdfs:label "Doroteo T. Toledano".} } \\ 
 -3699.0 & 1 & \makecell{\texttt{dsw:person/doroteo-t-toledano rdfs:label ?n1.} } \\ 
 -3699.2 & 1 & \makecell{\texttt{dsw:person/alberto-musetti ?p1 foaf:Person.} } \\ 
 -3699.2 & 2 & \makecell{\texttt{\_:genid152 rdf:\_1 ?n1.} } \\ 
 -3699.3 & 1 & \makecell{\texttt{dsw:person/michalis-foukarakis foaf:name ?n1.} } \\ 
 -3699.7 & 1 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 rdfs:label ?n1.} } \\ 
 -3699.8 & 1 & \makecell{\texttt{dsw:person/michalis-foukarakis rdfs:label ?n1.} } \\ 
 -3699.9 & 1 & \makecell{\texttt{dsw:person/alejandro-jaimes ?p1 foaf:Person.} } \\ 
 -3699.9 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/paper/in-use-industrial/138/authorlist rdf:\_1 ?n1.} } \\ 
 -3700.1 & 1 & \makecell{\texttt{dsw:person/alain-leger ?p1 foaf:Person.} } \\ 
 -3700.2 & 4 & \makecell{\texttt{\_:genid117 rdf:\_1 ?n1.} } \\ 
 -3700.4 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/nikos-loutas.} } \\ 
 -3700.5 & 4 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember swc:heldBy ?n1.} } \\ 
 -3700.8 & 1 & \makecell{\texttt{dsw:person/georg-lausen ?p1 foaf:Person.} } \\ 
 -3700.9 & 1 & \makecell{\texttt{?n1 rdfs:label "Graph Kernels for RDF data".} } \\ 
 -3700.9 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/petr-pollak.} } \\ 
 -3701.1 & 2 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/aidan-boran.} } \\ 
 -3701.8 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-633 rdfs:label ?n1.} } \\ 
 -3702.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/431/authorlist rdf:\_1 ?n1.} } \\ 
 -3702.1 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/bettina-bauer-messmer.} } \\ 
 -3702.1 & 2 & \makecell{\texttt{dsw:conference/www/2012/eu/10 swc:isPartOf ?n1.} } \\ 
 -3702.3 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/violeta-damjanovic.} } \\ 
 -3702.3 & 1 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 dc:creator ?n1.} } \\ 
 -3702.5 & 37 & \makecell{\texttt{?n1 ?p2 dsw:person/chiara-ghidini.} } \\ 
 -3702.5 & 1 & \makecell{\texttt{?n1 rdf:\_2 dsw:person/stephen-larson.} } \\ 
 -3702.8 & 2 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/siegfried-handschuh.} } \\ 
 -3702.9 & 1 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember rdf:type ?n1.} } \\ 
 -3702.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/468/authorlist rdf:\_1 ?n1.} } \\ 
 -3703.0 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/raluca-zaharia.} } \\ 
 -3703.4 & 4 & \makecell{\texttt{?n1 foaf:maker dsw:person/malte-kiesel.} } \\ 
 -3703.5 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/naimul-basher.} } \\ 
 -3703.6 & 37 & \makecell{\texttt{\_:genid94 ?p2 ?n1.} } \\ 
 -3703.8 & 4 & \makecell{\texttt{dsw:workshop/terra\_cognita/2008/paper/main/4 dc:creator ?n1.} } \\ 
 -3704.0 & 1 & \makecell{\texttt{?n1 dc:creator dsw:person/vinhtuan-thai.} } \\ 
 -3704.3 & 2 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/jerome-euzenat.} } \\ 
 -3704.7 & 39 & \makecell{\texttt{\_:genid62 ?p2 ?n1.} } \\ 
 -3705.4 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/tomas-vitvar.} } \\ 
 -3705.7 & 3 & \makecell{\texttt{?n1 swc:heldBy dsw:person/tomas-vitvar.} } \\ 
 -3706.6 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/aidan-boran.} } \\ 
 -3706.9 & 3 & \makecell{\texttt{?n1 swrs:author dsw:person/hiroyuki-shinnou.} } \\ 
 -3708.8 & 2 & \makecell{\texttt{\_:genid65 http://www.w3.org/2002/12/cal/ical\_tzid ?n1.} } \\ 
 -3709.0 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-766/authors rdf:type ?n1.} } \\ 
 -3710.4 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/achille-fokoue.} } \\ 
 -3710.8 & 2 & \makecell{\texttt{?n1 swc:heldBy dsw:person/aidan-boran.} } \\ 
 -3711.7 & 6 & \makecell{\texttt{?n1 ?p2 dsw:person/bruno-alves.} } \\ 
 -3714.2 & 1 & \makecell{\texttt{dsw:conference/eswc/2010/paper/services/3/authorlist ?p2 ?n1.} } \\ 
 -3714.3 & 5 & \makecell{\texttt{?n1 swc:heldBy dsw:person/chiara-ghidini.} } \\ 
 -3714.9 & 1 & \makecell{\texttt{dsw:person/georg-lausen rdf:type ?n1.} } \\ 
 -3715.9 & 12 & \makecell{\texttt{?n1 rdf:\_3 dsw:person/eero-hyvoenen.} } \\ 
 -3716.7 & 5 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/andreas-harth.} } \\ 
 -3717.0 & 9 & \makecell{\texttt{dsw:person/georg-lausen swc:holdsRole ?n1.} } \\ 
 -3717.4 & 13 & \makecell{\texttt{\_:genid38 rdf:\_1 ?n1.} \\\texttt{dsw:conference/iswc/2008/paper/poster\_demo/18 ?p2 \_:genid38.} } \\ 
 -3718.4 & 8 & \makecell{\texttt{?n1 swc:heldBy dsw:person/achille-fokoue.} } \\ 
 -3719.7 & 37 & \makecell{\texttt{\_:genid59 ?p2 ?n1.} } \\ 
 -3719.9 & 13 & \makecell{\texttt{?n1 ?p2 dsw:person/aidan-boran.} } \\ 
 -3720.5 & 20 & \makecell{\texttt{?n1 ?p2 dsw:person/tomas-vitvar.} } \\ 
 -3721.7 & 4 & \makecell{\texttt{dsw:conference/eswc/2012/paper/in-use-industrial/138/authorlist ?p2 ?n1.} } \\ 
 -3723.0 & 9 & \makecell{\texttt{?n1 dc:creator dsw:person/andreas-harth.} } \\ 
 -3724.0 & 12 & \makecell{\texttt{?n1 swc:heldBy dsw:person/bernhard-schandl.} } \\ 
 -3726.0 & 58 & \makecell{\texttt{dsw:conference/iswc/2009/paper/inuse/143 dc:creator dsw:person/achille-fokoue.} \\\texttt{?n1 ?p2 dsw:person/achille-fokoue.} } \\ 
 -3726.2 & 26 & \makecell{\texttt{?n1 swrs:author dsw:person/eero-hyvoenen.} } \\ 
 -3726.4 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/petr-pollak.} } \\ 
 -3727.7 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/benjamin-coe.} } \\ 
 -3728.1 & 24 & \makecell{\texttt{?n1 swc:heldBy dsw:person/mathieu-daquin.} } \\ 
 -3728.9 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/naimul-basher.} } \\ 
 -3729.6 & 0 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 rdf:type swc:Presenter.} } \\ 
 -3730.0 & 2 & \makecell{\texttt{dsw:conference/iswc/2010/paper/445/authorlist ?p2 ?n1.} } \\ 
 -3731.5 & 30 & \makecell{\texttt{\_:genid96 ?p2 ?n1.} } \\ 
 -3731.5 & 4 & \makecell{\texttt{?n1 ?p2 dsw:person/michael-pool.} } \\ 
 -3731.7 & 2 & \makecell{\texttt{dsw:conference/iswc/2010/paper/431/authorlist ?p2 ?n1.} } \\ 
 -3731.9 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/raluca-zaharia.} } \\ 
 -3732.4 & 13 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/18 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n1.} \\\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -3732.6 & 6 & \makecell{\texttt{dsw:workshop/ssws/2008/main/chair/organisingcommitteemember ?p2 ?n1.} } \\ 
 -3733.0 & 3 & \makecell{\texttt{?n2 rdfs:label "Graph Kernels for RDF data".} \\\texttt{?n2 dc:creator ?n1.} } \\ 
 -3735.3 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/julien-gaugaz.} } \\ 
 -3738.0 & 45 & \makecell{\texttt{http://www.eswc2007.org/pdf/eswc07-dellavalle.pdf rdf:type ?n2.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -3738.4 & 2 & \makecell{\texttt{?n1 ?p2 dsw:person/sukwon-chung.} } \\ 
 -3739.9 & 5 & \makecell{\texttt{\_:bNode14 ?p2 ?n1.} } \\ 
 -3740.6 & 7 & \makecell{\texttt{?n1 ?p2 dsw:person/bettina-bauer-messmer.} } \\ 
 -3741.2 & 5 & \makecell{\texttt{dsw:conference/iswc/2010/paper/489/authorlist ?p2 ?n1.} } \\ 
 -3741.4 & 6 & \makecell{\texttt{dsw:conference/iswc/2010/paper/261/authorlist ?p2 ?n1.} } \\ 
 -3742.5 & 36 & \makecell{\texttt{\_:genid99 ?p2 ?n1.} } \\ 
 -3743.8 & 16 & \makecell{\texttt{\_:genid121 ?p2 ?n1.} } \\ 
 -3744.2 & 0 & \makecell{\texttt{?n1 ?p4 ?n2.} \\\texttt{?n2 rdf:\_1 ?n3.} } \\ 
 -3746.9 & 4 & \makecell{\texttt{?n1 ?p2 dsw:person/stephen-larson.} } \\ 
 -3749.0 & 7 & \makecell{\texttt{dsw:conference/www/2012/demo/demo-chair ?p2 ?n1.} } \\ 
 -3749.7 & 20 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-633 ?p2 ?n1.} } \\ 
 -3750.6 & 13 & \makecell{\texttt{\_:genid152 ?p2 ?n1.} } \\ 
 -3751.7 & 29 & \makecell{\texttt{\_:genid65 ?p2 ?n1.} } \\ 
 -3752.6 & 12 & \makecell{\texttt{?n1 ?p2 dsw:person/vinhtuan-thai.} } \\ 
 -3753.1 & 6 & \makecell{\texttt{dsw:person/doroteo-t-toledano ?p2 ?n1.} } \\ 
 -3755.1 & 59 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/tracks/posters-demos/papers/375 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
 -3755.4 & 9 & \makecell{\texttt{?n1 ?p2 dsw:person/violeta-damjanovic.} } \\ 
 -3759.6 & 34 & \makecell{\texttt{\_:genid69 ?p2 ?n1.} } \\ 
 -3761.7 & 9 & \makecell{\texttt{dsw:person/alberto-musetti ?p2 ?n1.} } \\ 
 -3763.9 & 7 & \makecell{\texttt{dsw:person/michalis-foukarakis ?p2 ?n1.} } \\ 
 -3764.3 & 17 & \makecell{\texttt{\_:genid150 ?p2 ?n1.} } \\ 
 -3769.0 & 15 & \makecell{\texttt{?n1 ?p2 dsw:person/malte-kiesel.} } \\ 
 -3770.1 & 10 & \makecell{\texttt{dsw:conference/isemantics/2009/paper/main/11 ?p2 ?n1.} } \\ 
 -3772.9 & 45 & \makecell{\texttt{?n1 ?p2 dsw:person/elena-paslaru-bontas-simperl.} } \\ 
 -3780.1 & 18 & \makecell{\texttt{dsw:conference/iswc/2008/paper/poster\_demo/43 ?p2 ?n1.} } \\ 
 -3780.2 & 18 & \makecell{\texttt{dsw:conference/eswc/2006/demo-papers/FD01-hitzler ?p2 ?n1.} } \\ 
 -3784.0 & 1290 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 -3797.8 & 12 & \makecell{\texttt{dsw:person/alejandro-jaimes ?p2 ?n1.} } \\ 
 -3800.4 & 26 & \makecell{\texttt{dsw:person/georg-lausen ?p2 ?n1.} } \\ 
 -3833.3 & 233 & \makecell{\texttt{?n1 ?p5 ?n3.} \\\texttt{?n3 ?p4 ?n2.} } \\ 
 -3902.5 & 47 & \makecell{\texttt{?n1 rdf:\_10 ?n2.} } \\ 
 -3993.4 & 405 & \makecell{\texttt{?n1 ?p4 ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 -4021.2 & 2102 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 -4106.8 & 156 & \makecell{\texttt{?n1 rdfs:label ?n2.} } \\ 
 -4204.5 & 373 & \makecell{\texttt{?n1 ?p4 ?n3.} \\\texttt{?n2 ?p4 ?n3.} } \\ 
 -4750.2 & 2914 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -4761.6 & 248 & \makecell{\texttt{?n1 ?p2 foaf:Person.} } \\ 
 -4802.7 & 293 & \makecell{\texttt{?n1 rdf:type foaf:Person.} } \\ 
 -5632.4 & 301 & \makecell{\texttt{?n1 foaf:name ?n2.} } \\ 
 -6987.6 & 3047 & \makecell{\texttt{?n1 rdf:\_2 ?n2.} } \\ 
 -9429.5 & 2385 & \makecell{\texttt{?n1 foaf:maker ?n2.} } \\ 
 -10493.1 & 2528 & \makecell{\texttt{?n1 dc:creator ?n2.} } \\ 
\hline
\end{longtable}

\subsection{Dogfood, top 500 by frequency}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 -5052.5 & 16582 & \makecell{\texttt{?n1 rdf:\_1 ?n3.} \\\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -4021.9 & 6804 & \makecell{\texttt{\_:genid9 ?p4 ?n1.} \\\texttt{\_:genid9 ?p3 ?n2.} } \\ 
 28232.1 & 5807 & \makecell{\texttt{?n1 swc:heldBy ?n2.} } \\ 
 122602.4 & 5709 & \makecell{\texttt{?n1 swc:heldBy ?n2.} \\\texttt{?n2 swc:holdsRole ?n1.} } \\ 
 -3801.0 & 3734 & \makecell{\texttt{?n1 ?p3 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{?n2 ?p3 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -4178.4 & 3157 & \makecell{\texttt{?n2 ?p4 ?n1.} \\\texttt{?n2 rdf:\_4 ?n3.} } \\ 
 -7299.7 & 2799 & \makecell{\texttt{?n1 rdf:\_2 ?n2.} } \\ 
 -4442.9 & 2449 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -4599.6 & 2393 & \makecell{\texttt{?n1 rdf:\_1 ?n4.} \\\texttt{?n2 rdf:\_1 ?n4.} \\\texttt{?n2 rdf:\_1 ?n3.} } \\ 
 -9551.9 & 2193 & \makecell{\texttt{?n1 foaf:maker ?n2.} } \\ 
 6218.0 & 1939 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 -3804.8 & 1938 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/chair\_12 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
 -1305.2 & 1713 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator ?n2.} } \\ 
 -15966.3 & 1672 & \makecell{\texttt{?n1 http://purl.org/ontology/bibo/authorList ?n2.} } \\ 
 3849.2 & 1663 & \makecell{\texttt{?n1 swc:holdsRole ?n2.} } \\ 
 -1446.4 & 1587 & \makecell{\texttt{?n1 rdf:type rdf:Seq.} } \\ 
 271.9 & 1587 & \makecell{\texttt{?n1 ?p2 rdf:Seq.} } \\ 
 -2977.3 & 1541 & \makecell{\texttt{?n2 rdf:\_3 ?n3.} \\\texttt{?n2 http://www.w3.org/2002/12/cal/ical\_date ?n1.} } \\ 
 -197.2 & 1466 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} } \\ 
 -1595.5 & 1463 & \makecell{\texttt{?n2 rdf:type ?n3.} \\\texttt{?n2 rdf:\_4 ?n1.} } \\ 
 -3974.4 & 1292 & \makecell{\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
 -3829.0 & 1216 & \makecell{\texttt{\_:genid9 ?p3 ?n2.} \\\texttt{\_:genid9 ?p3 ?n1.} } \\ 
 -3828.3 & 1209 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/chair\_12 ?p5 ?n3.} \\\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
 -3082.3 & 1044 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -4012.3 & 670 & \makecell{\texttt{?n1 ?p5 ?n4.} \\\texttt{?n2 ?p5 ?n4.} \\\texttt{?n2 ?p5 ?n3.} } \\ 
 -4179.5 & 613 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 rdf:\_7 ?n3.} } \\ 
 -3790.2 & 479 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n1 ?p3 dsw:person/jason-jung.} } \\ 
 -3067.7 & 421 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -3808.7 & 398 & \makecell{\texttt{dsw:person/rayid-ghani ?p3 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1939.1 & 396 & \makecell{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date ?n2.} } \\ 
 -3745.9 & 359 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole ?n2.} \\\texttt{?n1 swc:holdsRole ?n2.} } \\ 
 -2779.2 & 335 & \makecell{\texttt{?n1 ?p2 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -2727.6 & 334 & \makecell{\texttt{?n1 swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 1626.1 & 334 & \makecell{\texttt{dsw:conference/eswc/2011/programme-committee-member swc:heldBy ?n1.} \\\texttt{?n1 swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3707.5 & 333 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{?n1 swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3729.5 & 333 & \makecell{\texttt{dsw:person/rayid-ghani ?p2 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{?n1 ?p2 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3171.2 & 309 & \makecell{\texttt{?n1 rdf:type geo:Point.} } \\ 
 -2738.5 & 277 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member ?p2 ?n1.} } \\ 
 -2678.0 & 274 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member swc:heldBy ?n1.} } \\ 
 -3887.4 & 271 & \makecell{\texttt{\_:genid95 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n3.} \\\texttt{?n1 ?p4 ?n2.} } \\ 
 -3754.4 & 164 & \makecell{\texttt{\_:genid9 ?p3 ?n1.} \\\texttt{\_:genid9 http://www.w3.org/2002/12/cal/ical\_date ?n2.} } \\ 
 -3817.4 & 147 & \makecell{\texttt{\_:genid95 rdf:type ?n3.} \\\texttt{?n1 rdf:type ?n3.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -3708.6 & 126 & \makecell{\texttt{?n2 ?p3 ?n1.} \\\texttt{?n2 http://www.w3.org/2002/12/cal/ical\_date "2009-05-31T18:00:00".} } \\ 
 -3550.9 & 97 & \makecell{\texttt{\_:genid10 ?p2 ?n1.} } \\ 
 -4294.2 & 95 & \makecell{\texttt{?n1 ?p2 foaf:Person.} } \\ 
 -3625.0 & 89 & \makecell{\texttt{\_:genid5 ?p2 ?n1.} } \\ 
 -3505.4 & 89 & \makecell{\texttt{dsw:organization/yahoo-inc ?p2 ?n1.} } \\ 
 -3698.7 & 86 & \makecell{\texttt{\_:genid6 ?p2 ?n1.} } \\ 
 -3633.7 & 86 & \makecell{\texttt{\_:genid11 ?p2 ?n1.} } \\ 
 -3757.9 & 82 & \makecell{\texttt{?n2 ?p3 ?n1.} \\\texttt{?n2 rdf:\_2 dsw:person/jan-hettenhausen.} } \\ 
 -3468.0 & 73 & \makecell{\texttt{dsw:organization/yahoo-inc foaf:member ?n1.} } \\ 
 -3815.9 & 72 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -3755.1 & 70 & \makecell{\texttt{dsw:workshop/om/2008/paper/main/21 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
 -3596.9 & 67 & \makecell{\texttt{?n1 ?p2 dsw:person/haofen-wang.} } \\ 
 -3660.9 & 67 & \makecell{\texttt{\_:genid18 ?p2 ?n1.} } \\ 
 -3721.2 & 66 & \makecell{\texttt{\_:genid18 ?p2 ?n1.} \\\texttt{\_:genid18 rdf:\_1 dsw:person/gary-wills.} } \\ 
 -3576.6 & 65 & \makecell{\texttt{?n1 ?p2 dsw:person/deborah-mcguinness.} } \\ 
 -3719.8 & 65 & \makecell{\texttt{\_:genid1 ?p2 ?n1.} } \\ 
 -3970.0 & 64 & \makecell{\texttt{?n1 rdf:\_9 ?n2.} } \\ 
 -3695.6 & 59 & \makecell{\texttt{\_:genid27 ?p2 ?n1.} } \\ 
 -3696.5 & 59 & \makecell{\texttt{\_:genid50 ?p2 ?n1.} } \\ 
 -3688.2 & 52 & \makecell{\texttt{\_:genid24 ?p2 ?n1.} } \\ 
 -3740.6 & 52 & \makecell{\texttt{?n1 ?p2 dsw:person/denny-vrandecic.} } \\ 
 -3737.8 & 51 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n2.} \\\texttt{?n1 swrs:author ?n2.} } \\ 
 -3679.5 & 49 & \makecell{\texttt{\_:genid23 ?p2 ?n1.} } \\ 
 -3706.1 & 48 & \makecell{\texttt{\_:genid48 ?p2 ?n1.} } \\ 
 -3716.6 & 48 & \makecell{\texttt{\_:genid48 ?p2 ?n1.} \\\texttt{dsw:conference/www/2008/paper/431 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid48.} } \\ 
 -3661.4 & 47 & \makecell{\texttt{?n1 ?p2 dsw:person/maria-esther-vidal.} } \\ 
 -3681.1 & 47 & \makecell{\texttt{\_:genid46 ?p2 ?n1.} } \\ 
 -3692.9 & 47 & \makecell{\texttt{\_:genid36 ?p2 ?n1.} } \\ 
 -3609.9 & 43 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member ?p2 ?n1.} } \\ 
 -3729.1 & 43 & \makecell{\texttt{\_:genid56 ?p2 ?n1.} } \\ 
 -3677.4 & 41 & \makecell{\texttt{\_:genid45 ?p2 ?n1.} } \\ 
 -3564.6 & 40 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member swc:heldBy ?n1.} } \\ 
 -3611.2 & 39 & \makecell{\texttt{?n1 ?p2 dsw:person/gong-cheng.} } \\ 
 -3698.4 & 39 & \makecell{\texttt{?n1 ?p2 dsw:person/matthew-rowe.} } \\ 
 -3743.1 & 38 & \makecell{\texttt{?n1 ?p2 dsw:person/giorgos-stoilos.} } \\ 
 -3725.6 & 37 & \makecell{\texttt{\_:genid82 ?p2 ?n1.} } \\ 
 -3597.5 & 37 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26 ?p2 ?n1.} } \\ 
 -3716.3 & 36 & \makecell{\texttt{?n1 ?p2 dsw:person/katharina-siorpaes.} } \\ 
 -3758.7 & 36 & \makecell{\texttt{\_:genid66 ?p2 ?n1.} } \\ 
 -3730.4 & 36 & \makecell{\texttt{\_:genid78 ?p2 ?n1.} } \\ 
 -3649.8 & 36 & \makecell{\texttt{?n1 ?p2 dsw:person/vanja-josifovski.} } \\ 
 -3759.6 & 34 & \makecell{\texttt{\_:genid69 ?p2 ?n1.} } \\ 
 -3731.8 & 33 & \makecell{\texttt{\_:genid72 ?p2 ?n1.} } \\ 
 -3693.2 & 32 & \makecell{\texttt{dsw:workshop/sdow/2011/pcmember ?p2 ?n1.} } \\ 
 -3745.8 & 31 & \makecell{\texttt{\_:genid90 ?p2 ?n1.} } \\ 
 -3732.2 & 30 & \makecell{\texttt{\_:genid9 rdf:\_3 ?n2.} \\\texttt{\_:genid9 http://www.w3.org/2002/12/cal/ical\_date ?n1.} } \\ 
 -3730.5 & 30 & \makecell{\texttt{?n1 rdf:\_1 ?n2.} \\\texttt{?n1 rdf:\_1 dsw:person/jason-jung.} } \\ 
 -3751.7 & 29 & \makecell{\texttt{\_:genid65 ?p2 ?n1.} } \\ 
 -3737.3 & 26 & \makecell{\texttt{?n1 ?p2 dsw:person/karl-aberer.} } \\ 
 -3665.6 & 24 & \makecell{\texttt{\_:genid5 rdf:\_1 ?n1.} } \\ 
 -3737.1 & 23 & \makecell{\texttt{?n1 ?p2 dsw:person/gang-wang.} } \\ 
 -3703.3 & 21 & \makecell{\texttt{?n1 rdf:\_3 ?n2.} \\\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2009-05-31T18:00:00".} } \\ 
 -3728.1 & 21 & \makecell{\texttt{?n1 ?p3 "2009-05-31T18:00:00".} \\\texttt{?n1 rdf:\_3 ?n2.} } \\ 
 -3769.4 & 20 & \makecell{\texttt{\_:genid135 ?p2 ?n1.} } \\ 
 -3670.9 & 19 & \makecell{\texttt{\_:genid15 rdf:\_1 ?n1.} } \\ 
 -3697.6 & 16 & \makecell{\texttt{\_:genid18 rdf:\_2 ?n1.} } \\ 
 -3733.0 & 15 & \makecell{\texttt{?n1 foaf:maker dsw:person/deborah-mcguinness.} } \\ 
 -3716.3 & 15 & \makecell{\texttt{?n1 ?p2 dsw:person/rolf-gruetter.} } \\ 
 -3787.8 & 14 & \makecell{\texttt{dsw:person/adrian-marte ?p2 ?n1.} } \\ 
 -3788.7 & 14 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 ?p2 ?n1.} } \\ 
 -3722.1 & 14 & \makecell{\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 ?p2 ?n1.} } \\ 
 -3674.6 & 14 & \makecell{\texttt{\_:genid23 rdf:\_2 ?n1.} } \\ 
 -3729.2 & 13 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p2 ?n1.} } \\ 
 -3691.8 & 13 & \makecell{\texttt{\_:genid48 rdf:\_1 ?n1.} } \\ 
 -3673.6 & 13 & \makecell{\texttt{?n1 ?p2 \_:genid48.} } \\ 
 -3646.4 & 13 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid48.} } \\ 
 -3690.6 & 13 & \makecell{\texttt{\_:genid27 rdf:\_3 ?n1.} } \\ 
 -3717.4 & 12 & \makecell{\texttt{?n1 ?p2 dsw:person/samur-araujo.} } \\ 
 -3736.6 & 11 & \makecell{\texttt{?n1 ?p2 dsw:person/kai-eckert.} } \\ 
 -3703.9 & 11 & \makecell{\texttt{?n1 ?p2 dsw:person/james-michaelis.} } \\ 
 -3717.9 & 10 & \makecell{\texttt{?n1 ?p2 dsw:person/varish-mulwad.} } \\ 
 -3786.8 & 9 & \makecell{\texttt{dsw:person/kenneth-church ?p2 ?n1.} } \\ 
 -3704.2 & 9 & \makecell{\texttt{?n1 ?p2 dsw:person/jianqiang-li.} } \\ 
 -3714.2 & 9 & \makecell{\texttt{?n1 swc:heldBy dsw:person/matthew-rowe.} } \\ 
 -3764.5 & 8 & \makecell{\texttt{dsw:person/nematollaah-shiri ?p2 ?n1.} } \\ 
 -3704.0 & 8 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator dsw:person/fabio-ciravegna.} } \\ 
 -3672.5 & 8 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid72.} } \\ 
 -3694.1 & 8 & \makecell{\texttt{\_:genid78 rdf:\_2 ?n1.} } \\ 
 -3703.7 & 7 & \makecell{\texttt{dsw:organization/yahoo-inc rdfs:label ?n1.} } \\ 
 -3699.9 & 7 & \makecell{\texttt{dsw:organization/yahoo-inc foaf:name ?n1.} } \\ 
 -3744.7 & 6 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors ?p2 ?n1.} } \\ 
 -3705.6 & 6 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/willem-van-hage.} } \\ 
 -3742.8 & 5 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515/authorlist ?p2 ?n1.} } \\ 
 -3726.0 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/lisa-lena-opas-haenninen.} } \\ 
 -3724.1 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/yi-chin-tu.} } \\ 
 -3729.5 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/yon-dourisboure.} } \\ 
 -3726.8 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/kukulenz-dirk.} } \\ 
 -3729.2 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/jan-zemanek.} } \\ 
 -3745.6 & 5 & \makecell{\texttt{\_:b5 ?p2 ?n1.} } \\ 
 -3694.8 & 5 & \makecell{\texttt{\_:genid72 rdf:\_3 ?n1.} } \\ 
 -3726.4 & 5 & \makecell{\texttt{?n1 ?p2 dsw:person/brett-cannon.} } \\ 
 -3711.8 & 5 & \makecell{\texttt{\_:genid1 rdf:\_1 ?n1.} } \\ 
 -3680.9 & 4 & \makecell{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} } \\ 
 -3705.2 & 4 & \makecell{\texttt{?n1 ?p2 "2010-05-31T13:00:00".} } \\ 
 -3663.1 & 4 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26 ?p1 dsw:person/james-michaelis.} } \\ 
 -3739.7 & 4 & \makecell{\texttt{dsw:conference/www/2012/paper/518/talk/speaker ?p2 ?n1.} } \\ 
 -3707.7 & 3 & \makecell{\texttt{\_:genid135 rdf:\_1 ?n1.} } \\ 
 -3725.5 & 3 & \makecell{\texttt{dsw:workshop/ores/2010/paper/main/9/authorlist ?p2 ?n1.} } \\ 
 -3703.2 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/gang-wang.} } \\ 
 -3703.3 & 3 & \makecell{\texttt{?n1 dc:creator dsw:person/samur-araujo.} } \\ 
 -3696.2 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/samur-araujo.} } \\ 
 -3667.3 & 3 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/68 ?p1 dsw:person/jianqiang-li.} } \\ 
 -3668.7 & 3 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 dsw:person/jianqiang-li.} } \\ 
 -3668.7 & 3 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 dsw:person/yu-zhao.} } \\ 
 -3682.9 & 3 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 dsw:person/jianqiang-li.} \\\texttt{dsw:workshop/LDOW/2008/paper/3 swc:isPartOf dsw:workshop/LDOW/2008/proceedings.} } \\ 
 -3684.5 & 3 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/varish-mulwad.} } \\ 
 -3711.5 & 3 & \makecell{\texttt{?n1 foaf:member dsw:person/xuanhui-wang.} } \\ 
 -3692.5 & 3 & \makecell{\texttt{?n1 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid132.} } \\ 
 -3718.2 & 3 & \makecell{\texttt{\_:genid69 rdf:type ?n1.} } \\ 
 -3707.6 & 3 & \makecell{\texttt{\_:genid18 ?p2 "2010-05-31T13:00:00".} \\\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} } \\ 
 -3684.5 & 3 & \makecell{\texttt{?n1 http://www.w3.org/2002/12/cal/ical\_date "2008-04-25T15:30:00".} } \\ 
 -3707.5 & 2 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole ?n1.} } \\ 
 -3704.5 & 2 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 foaf:maker ?n1.} } \\ 
 -3703.5 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/yu-zhao.} } \\ 
 -3703.5 & 2 & \makecell{\texttt{?n1 foaf:maker dsw:person/jianqiang-li.} } \\ 
 -3706.2 & 2 & \makecell{\texttt{?n1 swc:heldBy dsw:person/kai-eckert.} } \\ 
 -3694.9 & 2 & \makecell{\texttt{dsw:organization/yahoo-inc ?p1 "Yahoo! Inc".} } \\ 
 -3732.0 & 2 & \makecell{\texttt{?n1 ?p2 "Yahoo! Inc".} } \\ 
 -3712.6 & 2 & \makecell{\texttt{\_:genid11 ?p2 ?n1.} \\\texttt{dsw:workshop/obi/2008/paper/main/7 swrs:author ?n1.} } \\ 
 -3706.3 & 2 & \makecell{\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 swrs:author ?n1.} } \\ 
 -3701.4 & 2 & \makecell{\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 http://purl.org/dc/terms/creator ?n1.} } \\ 
 -3708.3 & 2 & \makecell{\texttt{\_:genid10 http://www.w3.org/2002/12/cal/ical\_tzid ?n1.} } \\ 
 -3703.8 & 1 & \makecell{\texttt{dsw:person/kenneth-church swc:holdsRole ?n1.} } \\ 
 -3689.9 & 1 & \makecell{\texttt{dsw:person/lorenzino-vaccari ?p1 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{dsw:person/rayid-ghani ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3667.9 & 1 & \makecell{\texttt{dsw:person/lorenzino-vaccari swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{dsw:person/rayid-ghani swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3692.5 & 1 & \makecell{\texttt{dsw:person/lorenzino-vaccari ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3691.0 & 1 & \makecell{\texttt{dsw:person/neil-ireson ?p1 dsw:conference/eswc/2011/programme-committee-member.} \\\texttt{dsw:person/rayid-ghani ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3692.7 & 1 & \makecell{\texttt{dsw:person/rayid-ghani ?p1 dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3670.7 & 1 & \makecell{\texttt{dsw:person/rayid-ghani swc:holdsRole dsw:conference/eswc/2011/programme-committee-member.} } \\ 
 -3669.6 & 1 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member swc:heldBy dsw:person/maria-esther-vidal.} } \\ 
 -3691.6 & 1 & \makecell{\texttt{dsw:workshop/swcs/2012/programme-committee-member ?p1 dsw:person/maria-esther-vidal.} } \\ 
 -3688.4 & 1 & \makecell{\texttt{dsw:organization/universidad-simon-bolivar ?p1 dsw:person/maria-esther-vidal.} } \\ 
 -3719.8 & 1 & \makecell{\texttt{dsw:workshop/ores/2010/paper/main/9/authorlist ?p2 ?n1.} \\\texttt{?n1 foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3697.7 & 1 & \makecell{\texttt{dsw:workshop/ores/2010/paper/main/9/authorlist rdf:\_3 ?n1.} \\\texttt{?n1 foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3707.2 & 1 & \makecell{\texttt{?n1 foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3671.1 & 1 & \makecell{\texttt{dsw:person/adrian-marte foaf:mbox\_sha1sum "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3729.2 & 1 & \makecell{\texttt{?n1 ?p2 "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3693.1 & 1 & \makecell{\texttt{dsw:person/adrian-marte ?p1 "7ae854a556d73b4957e15b6f13efc08eaa48e3e4".} } \\ 
 -3685.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515/authorlist ?p1 dsw:person/samur-araujo.} } \\ 
 -3663.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515/authorlist rdf:\_1 dsw:person/samur-araujo.} } \\ 
 -3680.4 & 1 & \makecell{\texttt{dsw:conference/iswc/2010/paper/515 foaf:maker dsw:person/samur-araujo.} \\\texttt{dsw:conference/iswc/2010/paper/515/authorlist ?p1 dsw:person/samur-araujo.} } \\ 
 -3701.3 & 1 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator dsw:person/jianqiang-li.} } \\ 
 -3665.1 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/68 http://purl.org/dc/terms/creator dsw:person/jianqiang-li.} } \\ 
 -3666.5 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 foaf:maker dsw:person/jianqiang-li.} } \\ 
 -3666.5 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 foaf:maker dsw:person/yu-zhao.} } \\ 
 -3690.7 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 ?p1 http://sunsite.informatik.rwth-aachen.de/Publications/CEUR-WS/Vol-369/paper04.pdf.} } \\ 
 -3666.3 & 1 & \makecell{\texttt{dsw:workshop/LDOW/2008/paper/3 dc:creator dsw:person/jianqiang-li.} } \\ 
 -3699.5 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/lisa-lena-opas-haenninen.} } \\ 
 -3663.3 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-775/authors rdf:\_1 dsw:person/lisa-lena-opas-haenninen.} } \\ 
 -3685.3 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-775/authors ?p1 dsw:person/lisa-lena-opas-haenninen.} } \\ 
 -3705.0 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri foaf:firstName ?n1.} } \\ 
 -3698.6 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri ?p1 foaf:Person.} } \\ 
 -3676.6 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri rdf:type foaf:Person.} } \\ 
 -3712.8 & 1 & \makecell{\texttt{dsw:person/nematollaah-shiri rdf:type ?n1.} } \\ 
 -3686.6 & 1 & \makecell{\texttt{\_:genid24 ?p1 dsw:person/yi-chin-tu.} } \\ 
 -3731.0 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 http://purl.org/ontology/bibo/authorList ?n1.} \\\texttt{?n1 rdf:\_1 ?n2.} } \\ 
 -3705.5 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 http://purl.org/ontology/bibo/authorList ?n1.} } \\ 
 -3691.4 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 ?p1 dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
 -3727.5 & 1 & \makecell{\texttt{?n1 ?p2 dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
 -3669.3 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/poster/ab-643 http://purl.org/ontology/bibo/authorList dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
 -3705.5 & 1 & \makecell{\texttt{?n1 http://purl.org/ontology/bibo/authorList dsw:conference/dh/2010/abstracts/poster/ab-643/authors.} } \\ 
 -3670.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member swc:heldBy dsw:person/kai-eckert.} } \\ 
 -3692.1 & 1 & \makecell{\texttt{dsw:conference/eswc/2012/research-track-committee-member ?p1 dsw:person/kai-eckert.} } \\ 
 -3750.6 & 1 & \makecell{\texttt{?n1 ?p3 ?n2.} \\\texttt{?n2 foaf:made dsw:conference/www/2010/paper/main/23.} } \\ 
 -3688.0 & 1 & \makecell{\texttt{\_:genid90 ?p1 dsw:person/jan-zemanek.} } \\ 
 -3675.7 & 1 & \makecell{\texttt{\_:genid15 rdf:\_1 dsw:person/c-maria-keet.} } \\ 
 -3671.7 & 1 & \makecell{\texttt{\_:genid15 rdf:\_1 dsw:person/daniel-bachlechner.} } \\ 
 -3699.6 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors rdf:\_1 ?n1.} } \\ 
 -3702.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors rdf:\_2 ?n1.} } \\ 
 -3666.6 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors rdf:\_2 dsw:person/deborah-mcguinness.} } \\ 
 -3688.7 & 1 & \makecell{\texttt{dsw:conference/iswc/2012/paper/poster-demo-26/authors ?p1 dsw:person/deborah-mcguinness.} } \\ 
 -3712.0 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/doctoral-consortium/14/authorlist ?p2 ?n1.} } \\ 
 -3675.9 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/doctoral-consortium/14/authorlist ?p1 dsw:person/varish-mulwad.} } \\ 
 -3653.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/doctoral-consortium/14/authorlist rdf:\_1 dsw:person/varish-mulwad.} } \\ 
 -3662.5 & 1 & \makecell{\texttt{\_:b3 rdf:\_1 dsw:person/varish-mulwad.} } \\ 
 -3684.5 & 1 & \makecell{\texttt{\_:b3 ?p1 dsw:person/varish-mulwad.} } \\ 
 -3698.7 & 1 & \makecell{\texttt{\_:b3 rdf:\_1 ?n1.} } \\ 
 -3667.4 & 1 & \makecell{\texttt{dsw:organization/yahoo-inc rdfs:label "Yahoo! Inc".} } \\ 
 -3670.4 & 1 & \makecell{\texttt{dsw:organization/yahoo-inc foaf:member dsw:person/xuanhui-wang.} } \\ 
 -3695.7 & 1 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n1.} \\\texttt{dsw:workshop/obi/2008/paper/main/7 swrs:author ?n1.} } \\ 
 -3699.7 & 1 & \makecell{\texttt{\_:genid11 rdf:\_1 ?n1.} \\\texttt{dsw:conference/iswc/2011/paper/poster-demo/40 swrs:author ?n1.} } \\ 
 -3698.8 & 1 & \makecell{\texttt{?n1 http://purl.org/dc/terms/creator dsw:person/akira-mori.} } \\ 
 -3699.2 & 1 & \makecell{\texttt{?n1 swrs:editor dsw:person/denny-vrandecic.} } \\ 
 -3663.1 & 1 & \makecell{\texttt{dsw:workshop/insemtive/2008/proceedings swrs:editor dsw:person/denny-vrandecic.} } \\ 
 -3685.1 & 1 & \makecell{\texttt{dsw:workshop/insemtive/2008/proceedings ?p1 dsw:person/denny-vrandecic.} } \\ 
 -3690.3 & 1 & \makecell{\texttt{\_:genid48 ?p1 dsw:person/willem-van-hage.} } \\ 
 -3668.3 & 1 & \makecell{\texttt{\_:genid48 rdf:\_1 dsw:person/willem-van-hage.} } \\ 
 -3665.3 & 1 & \makecell{\texttt{dsw:conference/www/2008/paper/431 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid48.} } \\ 
 -3687.4 & 1 & \makecell{\texttt{dsw:conference/www/2008/paper/431 ?p1 \_:genid48.} } \\ 
 -3701.7 & 1 & \makecell{\texttt{dsw:conference/www/2007/paper/main/429 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n1.} } \\ 
 -3703.1 & 1 & \makecell{\texttt{?n1 foaf:maker dsw:person/brett-cannon.} } \\ 
 -3691.2 & 1 & \makecell{\texttt{\_:genid36 ?p1 dsw:person/vanja-josifovski.} } \\ 
 -3728.0 & 1 & \makecell{\texttt{dsw:workshop/om/2008/paper/main/21 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList ?n1.} \\\texttt{?n1 rdf:\_7 ?n2.} } \\ 
 -3693.5 & 1 & \makecell{\texttt{\_:genid16 rdf:\_7 ?n1.} \\\texttt{dsw:workshop/om/2008/paper/main/21 vu:~mcaklein/onto/swrc\_ext/2005/05\_authorList \_:genid16.} } \\ 
 -3690.1 & 1 & \makecell{\texttt{\_:genid23 ?p1 dsw:person/zheng-chen.} } \\ 
 -3697.6 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/semantic-web-in-use/76/authorlist rdf:\_9 ?n1.} } \\ 
 -3691.1 & 1 & \makecell{\texttt{dsw:workshop/sdow/2011/pcmember ?p1 dsw:person/matthew-rowe.} } \\ 
 -3669.0 & 1 & \makecell{\texttt{dsw:workshop/sdow/2011/pcmember swc:heldBy dsw:person/matthew-rowe.} } \\ 
 -3690.8 & 1 & \makecell{\texttt{dsw:conference/iswc/2011/paper/industry/21/authorlist rdf:\_1 ?n1.} } \\ 
 -3707.0 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/karl-aberer.} } \\ 
 -3670.8 & 1 & \makecell{\texttt{\_:genid1 rdf:\_1 dsw:person/karl-aberer.} } \\ 
 -3692.9 & 1 & \makecell{\texttt{\_:genid1 ?p1 dsw:person/karl-aberer.} } \\ 
 -3700.1 & 1 & \makecell{\texttt{?n1 rdf:\_1 dsw:person/peyman-sazedj.} } \\ 
 -3700.1 & 1 & \makecell{\texttt{\_:genid18 http://www.w3.org/2002/12/cal/ical\_date ?n1.} } \\ 
 -3663.9 & 1 & \makecell{\texttt{\_:genid18 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} } \\ 
 -3686.0 & 1 & \makecell{\texttt{\_:genid18 ?p1 "2010-05-31T13:00:00".} } \\ 
 -3671.5 & 1 & \makecell{\texttt{\_:genid11 http://www.w3.org/2002/12/cal/ical\_date "2010-05-31T13:00:00".} \\\texttt{\_:genid18 ?p1 "2010-05-31T13:00:00".} } \\ 
 -3707.6 & 1 & \makecell{\texttt{\_:genid11 http://www.w3.org/2002/12/cal/ical\_date ?n1.} \\\texttt{\_:genid18 ?p2 ?n1.} } \\ 
 -3693.5 & 1 & \makecell{\texttt{\_:genid11 ?p2 "2010-05-31T13:00:00".} \\\texttt{\_:genid18 ?p1 "2010-05-31T13:00:00".} } \\ 
 -3809.4 & 1 & \makecell{\texttt{dsw:conference/iswc-aswc/2007/chair\_12 swc:heldBy ?n1.} \\\texttt{?n2 ?p4 ?n3.} \\\texttt{?n2 ?p4 ?n1.} } \\ 
 -3707.5 & 1 & \makecell{\texttt{dsw:conference/eswc/2007/time-20070605T153000EuropeVienna rdf:type ?n1.} } \\ 
 -3695.5 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors ?p1 rdf:Seq.} } \\ 
 -3673.4 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type rdf:Seq.} } \\ 
 -3709.6 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type ?n1.} } \\ 
 -3699.6 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type ?n1.} \\\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:\_4 dsw:person/stefan-beyer.} } \\ 
 -3735.8 & 1 & \makecell{\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:type ?n2.} \\\texttt{dsw:conference/dh/2010/abstracts/paper/ab-684/authors rdf:\_4 ?n1.} } \\ 
 -3709.7 & 1 & \makecell{\texttt{dsw:conference/www/2012/paper/518/talk/speaker rdf:type ?n1.} } \\ 
 -3744.2 & 0 & \makecell{\texttt{?n1 ?p4 ?n2.} \\\texttt{?n2 rdf:\_7 ?n3.} } \\ 
\hline
\end{longtable}

\subsection{AIFB, top 100 by log factor}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 8919.0 & 2477 & \makecell{\texttt{?n1 swrs:isAbout ?n2.} } \\ 
 3857.8 & 4124 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 923.9 & 1066 & \makecell{\texttt{?n1 ?p2 swrs:Person.} } \\ 
 827.3 & 1045 & \makecell{\texttt{?n1 rdf:type swrs:Person.} } \\ 
 -141.5 & 1240 & \makecell{\texttt{?n1 ?p2 swrs:Publication.} } \\ 
 -167.7 & 1222 & \makecell{\texttt{?n1 rdf:type swrs:Publication.} } \\ 
 -893.7 & 1529 & \makecell{\texttt{?n1 swrs:author ?n2.} } \\ 
 -907.8 & 3388 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -939.9 & 687 & \makecell{\texttt{?n1 rdf:type swrs:InProceedings.} } \\ 
 -975.2 & 687 & \makecell{\texttt{?n1 ?p2 swrs:InProceedings.} } \\ 
 -1495.3 & 31 & \makecell{\texttt{?n1 ?p2 owl:DatatypeProperty.} } \\ 
 -1536.6 & 146 & \makecell{\texttt{?n1 rdf:type swrs:ResearchTopic.} } \\ 
 -1582.5 & 152 & \makecell{\texttt{?n1 rdf:type owl:Restriction.} } \\ 
 -1583.0 & 147 & \makecell{\texttt{?n1 ?p2 swrs:ResearchTopic.} } \\ 
 -1609.0 & 152 & \makecell{\texttt{?n1 owl:onProperty ?n2.} } \\ 
 -1613.5 & 152 & \makecell{\texttt{?n1 ?p2 owl:Restriction.} } \\ 
 -1668.8 & 55 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
 -1696.9 & 55 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
 -1745.2 & 65 & \makecell{\texttt{?n1 rdf:type swrs:PhDStudent.} } \\ 
 -1747.9 & 40 & \makecell{\texttt{?n1 swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1767.8 & 78 & \makecell{\texttt{?n1 rdf:type swrs:Project.} } \\ 
 -1769.7 & 13 & \makecell{\texttt{swrs:InBook rdfs:subClassOf ?n1.} } \\ 
 -1775.0 & 87 & \makecell{\texttt{?n1 ?p2 swrs:Project.} } \\ 
 -1775.1 & 40 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1777.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1779.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1780.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type swrs:ResearchTopic.} } \\ 
 -1781.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1784.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Article.} } \\ 
 -1784.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance rdf:type swrs:Person.} } \\ 
 -1784.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance rdf:type swrs:Person.} } \\ 
 -1785.0 & 1 & \makecell{\texttt{\_:ub1bL199C18 rdf:type owl:Restriction.} } \\ 
 -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance rdf:type swrs:Person.} } \\ 
 -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id637instance rdf:type swrs:Person.} } \\ 
 -1785.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance rdf:type swrs:InProceedings.} } \\ 
 -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id684instance rdf:type swrs:InProceedings.} } \\ 
 -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type swrs:InProceedings.} } \\ 
 -1786.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance rdf:type swrs:InProceedings.} } \\ 
 -1787.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance rdf:type swrs:Person.} } \\ 
 -1787.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Publication.} } \\ 
 -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance rdf:type swrs:InProceedings.} } \\ 
 -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type swrs:Person.} } \\ 
 -1787.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type swrs:Publication.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2118instance rdf:type swrs:Person.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type swrs:Publication.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id709instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id703instance rdf:type swrs:Person.} } \\ 
 -1791.2 & 1 & \makecell{\texttt{swrs:title ?p1 owl:DatatypeProperty.} } \\ 
 -1796.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1796.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1798.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1799.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p1 swrs:ResearchTopic.} } \\ 
 -1800.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1800.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id877instance.} } \\ 
 -1801.0 & 14 & \makecell{\texttt{swrs:InBook ?p2 ?n1.} } \\ 
 -1802.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id28instance ?p1 swrs:ResearchTopic.} } \\ 
 -1802.2 & 227 & \makecell{\texttt{?n1 swrs:phone ?n2.} } \\ 
 -1803.3 & 1 & \makecell{\texttt{swrs:Proceedings ?p1 owl:Class.} } \\ 
 -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance ?p1 swrs:Person.} } \\ 
 -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id350instance ?p1 swrs:Person.} } \\ 
 -1803.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance ?p1 swrs:Person.} } \\ 
 -1804.0 & 1 & \makecell{\texttt{\_:ub1bL474C18 ?p1 owl:Restriction.} } \\ 
 -1804.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance ?p1 swrs:Person.} } \\ 
 -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id810instance ?p1 swrs:Person.} } \\ 
 -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id184instance ?p1 swrs:Person.} } \\ 
 -1805.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance ?p1 swrs:InProceedings.} } \\ 
 -1805.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p1 swrs:InProceedings.} } \\ 
 -1805.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance ?p1 swrs:InProceedings.} } \\ 
 -1806.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance ?p1 swrs:Person.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p1 swrs:Publication.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance ?p1 swrs:InProceedings.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p1 swrs:Person.} } \\ 
 -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance ?p1 swrs:Publication.} } \\ 
 -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id773instance ?p1 swrs:Publication.} } \\ 
 -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance ?p1 swrs:Publication.} } \\ 
 -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p1 swrs:Person.} } \\ 
 -1806.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id954instance ?p1 swrs:Publication.} } \\ 
 -1806.7 & 7 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication ?n1.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id820instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id786instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id848instance ?p1 swrs:Person.} } \\ 
 -1807.3 & 1 & \makecell{\texttt{\_:ub1bL560C18 ?p1 swrs:Person.} } \\ 
 -1808.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance swrs:name ?n1.} } \\ 
 -1808.9 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:worksAtProject ?n1.} } \\ 
 -1811.1 & 1 & \makecell{\texttt{?n1 swrs:name "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1811.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type ?n1.} } \\ 
 -1812.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name ?n1.} } \\ 
 -1812.1 & 1 & \makecell{\texttt{?n1 swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1813.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id68instance swrs:phone ?n1.} } \\ 
 -1813.3 & 9 & \makecell{\texttt{swrs:Organization ?p2 ?n1.} } \\ 
 -1813.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id79instance swrs:name ?n1.} } \\ 
 -1814.1 & 1 & \makecell{\texttt{\_:ub1bL474C18 owl:onProperty ?n1.} } \\ 
 -1814.8 & 1 & \makecell{\texttt{swrs:Organization rdf:type ?n1.} } \\ 
 -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id18instance rdf:type ?n1.} } \\ 
 -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id814instance rdf:type ?n1.} } \\ 
 -1815.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id20instance rdf:type ?n1.} } \\ 
 -1816.0 & 1 & \makecell{\texttt{\_:ub1bL422C18 rdf:type ?n1.} } \\ 
 -1816.0 & 1 & \makecell{\texttt{\_:ub1bL355C18 rdf:type ?n1.} } \\ 
 -1818.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type ?n1.} } \\ 
 -1818.8 & 152 & \makecell{\texttt{?n1 swrs:author http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id20instance.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id593instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id439instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id305instance rdf:type ?n1.} } \\ 
 -1820.7 & 22 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p2 ?n1.} } \\ 
 -1821.4 & 1 & \makecell{\texttt{swrs:member ?p2 ?n1.} } \\ 
 -1821.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n1.} } \\ 
 -1822.2 & 1 & \makecell{\texttt{swrs:name ?p2 ?n1.} } \\ 
 -1822.2 & 1 & \makecell{\texttt{swrs:title ?p2 ?n1.} } \\ 
 -1824.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id19instance rdf:type ?n1.} } \\ 
 -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type ?n1.} } \\ 
 -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1089instance rdf:type ?n1.} } \\ 
 -1825.1 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance rdf:type ?n1.} } \\ 
 -1825.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id552instance swrs:author ?n1.} } \\ 
 -1826.2 & 3 & \makecell{\texttt{\_:ub1bL658C27 ?p2 ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id617instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id516instance rdf:type ?n1.} } \\ 
 -1826.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type ?n1.} } \\ 
 -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1333instance rdf:type ?n1.} } \\ 
 -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id512instance rdf:type ?n1.} } \\ 
 -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance rdf:type ?n1.} } \\ 
 -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n1.} } \\ 
 -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id64instance rdf:type ?n1.} } \\ 
 -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id777instance rdf:type ?n1.} } \\ 
 -1827.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1306instance rdf:type ?n1.} } \\ 
 -1827.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type ?n1.} } \\ 
 -1828.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1166instance rdf:type ?n1.} } \\ 
 -1830.1 & 1 & \makecell{\texttt{?n1 ?p2 "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1831.1 & 1 & \makecell{\texttt{?n1 ?p2 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1833.9 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout ?n1.} } \\ 
 -1834.8 & 3 & \makecell{\texttt{?n1 ?p2 swrs:Graduate.} } \\ 
 -1837.2 & 3 & \makecell{\texttt{\_:ub1bL58C27 ?p2 ?n1.} } \\ 
 -1837.9 & 3 & \makecell{\texttt{\_:ub1bL474C18 ?p2 ?n1.} } \\ 
 -1839.0 & 3 & \makecell{\texttt{\_:ub1bL102C18 ?p2 ?n1.} } \\ 
 -1839.6 & 8 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p2 ?n1.} } \\ 
 -1839.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p2 ?n1.} } \\ 
 -1840.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id144instance ?p2 ?n1.} } \\ 
 -1840.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id190instance ?p2 ?n1.} } \\ 
 -1841.0 & 14 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Projekte/viewProjektOWL/id51instance swrs:member ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p2 ?n1.} } \\ 
 -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p2 ?n1.} } \\ 
 -1841.7 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id774instance ?p2 ?n1.} } \\ 
 -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p2 ?n1.} } \\ 
 -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id607instance ?p2 ?n1.} } \\ 
 -1843.9 & 6 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id530instance ?p2 ?n1.} } \\ 
 -1844.5 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance ?p2 ?n1.} } \\ 
 -1850.2 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance ?p2 ?n1.} } \\ 
 -1850.6 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2067instance ?p2 ?n1.} } \\ 
 -1860.0 & 1907 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n2.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -1869.8 & 17 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p2 ?n1.} } \\ 
 -1877.4 & 10 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1882.2 & 9 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p2 ?n1.} } \\ 
 -1883.4 & 18 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1272instance ?p2 ?n1.} } \\ 
 -1883.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p2 ?n1.} } \\ 
 -1887.6 & 1130 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1888.2 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id228instance ?p2 ?n1.} } \\ 
 -1895.2 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance ?p2 ?n1.} } \\ 
 -1903.8 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1305instance ?p2 ?n1.} } \\ 
 -1904.3 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1317instance ?p2 ?n1.} } \\ 
 -1908.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1152instance ?p2 ?n1.} } \\ 
 -1930.3 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1124instance ?p2 ?n1.} } \\ 
 -1931.5 & 2184 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance ?p3 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1937.4 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id882instance ?p2 ?n1.} } \\ 
 -1942.5 & 25 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p2 ?n1.} } \\ 
 -1942.7 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1025instance ?p2 ?n1.} } \\ 
 -2254.8 & 24274 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 -4014.6 & 494 & \makecell{\texttt{?n1 swrs:name ?n2.} } \\ 
\hline
\end{longtable}

\subsection{AIFB, top 100 by frequency}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 -2254.8 & 24274 & \makecell{\texttt{?n1 rdf:type ?n3.} \\\texttt{?n2 rdf:type ?n3.} } \\ 
 3857.8 & 4124 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 -907.8 & 3388 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 8919.0 & 2477 & \makecell{\texttt{?n1 swrs:isAbout ?n2.} } \\ 
 -1931.5 & 2184 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance ?p3 ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1860.0 & 1907 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n2.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -893.7 & 1529 & \makecell{\texttt{?n1 swrs:author ?n2.} } \\ 
 -141.5 & 1240 & \makecell{\texttt{?n1 ?p2 swrs:Publication.} } \\ 
 -167.7 & 1222 & \makecell{\texttt{?n1 rdf:type swrs:Publication.} } \\ 
 -1887.6 & 1130 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n2.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 923.9 & 1066 & \makecell{\texttt{?n1 ?p2 swrs:Person.} } \\ 
 827.3 & 1045 & \makecell{\texttt{?n1 rdf:type swrs:Person.} } \\ 
 -975.2 & 687 & \makecell{\texttt{?n1 ?p2 swrs:InProceedings.} } \\ 
 -939.9 & 687 & \makecell{\texttt{?n1 rdf:type swrs:InProceedings.} } \\ 
 -4014.6 & 494 & \makecell{\texttt{?n1 swrs:name ?n2.} } \\ 
 -1802.2 & 227 & \makecell{\texttt{?n1 swrs:phone ?n2.} } \\ 
 -1613.5 & 152 & \makecell{\texttt{?n1 ?p2 owl:Restriction.} } \\ 
 -1582.5 & 152 & \makecell{\texttt{?n1 rdf:type owl:Restriction.} } \\ 
 -1609.0 & 152 & \makecell{\texttt{?n1 owl:onProperty ?n2.} } \\ 
 -1818.8 & 152 & \makecell{\texttt{?n1 swrs:author http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id20instance.} } \\ 
 -1583.0 & 147 & \makecell{\texttt{?n1 ?p2 swrs:ResearchTopic.} } \\ 
 -1536.6 & 146 & \makecell{\texttt{?n1 rdf:type swrs:ResearchTopic.} } \\ 
 -1775.0 & 87 & \makecell{\texttt{?n1 ?p2 swrs:Project.} } \\ 
 -1767.8 & 78 & \makecell{\texttt{?n1 rdf:type swrs:Project.} } \\ 
 -1745.2 & 65 & \makecell{\texttt{?n1 rdf:type swrs:PhDStudent.} } \\ 
 -1668.8 & 55 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
 -1696.9 & 55 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
 -1775.1 & 40 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1747.9 & 40 & \makecell{\texttt{?n1 swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1495.3 & 31 & \makecell{\texttt{?n1 ?p2 owl:DatatypeProperty.} } \\ 
 -1942.5 & 25 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p2 ?n1.} } \\ 
 -1820.7 & 22 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p2 ?n1.} } \\ 
 -1937.4 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id882instance ?p2 ?n1.} } \\ 
 -1942.7 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1025instance ?p2 ?n1.} } \\ 
 -1930.3 & 20 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1124instance ?p2 ?n1.} } \\ 
 -1883.4 & 18 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1272instance ?p2 ?n1.} } \\ 
 -1869.8 & 17 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p2 ?n1.} } \\ 
 -1903.8 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1305instance ?p2 ?n1.} } \\ 
 -1904.3 & 16 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1317instance ?p2 ?n1.} } \\ 
 -1801.0 & 14 & \makecell{\texttt{swrs:InBook ?p2 ?n1.} } \\ 
 -1841.0 & 14 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Projekte/viewProjektOWL/id51instance swrs:member ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -1769.7 & 13 & \makecell{\texttt{swrs:InBook rdfs:subClassOf ?n1.} } \\ 
 -1908.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1152instance ?p2 ?n1.} } \\ 
 -1883.7 & 13 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p2 ?n1.} } \\ 
 -1888.2 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id228instance ?p2 ?n1.} } \\ 
 -1850.6 & 12 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2067instance ?p2 ?n1.} } \\ 
 -1833.9 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout ?n1.} } \\ 
 -1895.2 & 11 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance ?p2 ?n1.} } \\ 
 -1877.4 & 10 & \makecell{\texttt{?n1 ?p2 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1882.2 & 9 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p2 ?n1.} } \\ 
 -1813.3 & 9 & \makecell{\texttt{swrs:Organization ?p2 ?n1.} } \\ 
 -1839.6 & 8 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p2 ?n1.} } \\ 
 -1806.7 & 7 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication ?n1.} } \\ 
 -1843.9 & 6 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id530instance ?p2 ?n1.} } \\ 
 -1839.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p2 ?n1.} } \\ 
 -1850.2 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance ?p2 ?n1.} } \\ 
 -1825.9 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id552instance swrs:author ?n1.} } \\ 
 -1841.7 & 4 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id774instance ?p2 ?n1.} } \\ 
 -1837.9 & 3 & \makecell{\texttt{\_:ub1bL474C18 ?p2 ?n1.} } \\ 
 -1837.2 & 3 & \makecell{\texttt{\_:ub1bL58C27 ?p2 ?n1.} } \\ 
 -1840.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id190instance ?p2 ?n1.} } \\ 
 -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p2 ?n1.} } \\ 
 -1834.8 & 3 & \makecell{\texttt{?n1 ?p2 swrs:Graduate.} } \\ 
 -1841.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p2 ?n1.} } \\ 
 -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p2 ?n1.} } \\ 
 -1839.0 & 3 & \makecell{\texttt{\_:ub1bL102C18 ?p2 ?n1.} } \\ 
 -1844.5 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance ?p2 ?n1.} } \\ 
 -1826.2 & 3 & \makecell{\texttt{\_:ub1bL658C27 ?p2 ?n1.} } \\ 
 -1840.0 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id144instance ?p2 ?n1.} } \\ 
 -1842.8 & 3 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id607instance ?p2 ?n1.} } \\ 
 -1808.9 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:worksAtProject ?n1.} } \\ 
 -1821.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type ?n1.} } \\ 
 -1825.1 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id617instance rdf:type ?n1.} } \\ 
 -1827.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1306instance rdf:type ?n1.} } \\ 
 -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id512instance rdf:type ?n1.} } \\ 
 -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id572instance rdf:type ?n1.} } \\ 
 -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type ?n1.} } \\ 
 -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id64instance rdf:type ?n1.} } \\ 
 -1827.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type ?n1.} } \\ 
 -1827.0 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id777instance rdf:type ?n1.} } \\ 
 -1826.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1333instance rdf:type ?n1.} } \\ 
 -1824.6 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1089instance rdf:type ?n1.} } \\ 
 -1826.2 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id516instance rdf:type ?n1.} } \\ 
 -1824.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id19instance rdf:type ?n1.} } \\ 
 -1826.5 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type ?n1.} } \\ 
 -1828.7 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1166instance rdf:type ?n1.} } \\ 
 -1826.8 & 2 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id281instance rdf:type ?n1.} } \\ 
 -1798.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1779.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2115instance swrs:publication http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1202instance.} } \\ 
 -1804.0 & 1 & \makecell{\texttt{\_:ub1bL474C18 ?p1 owl:Restriction.} } \\ 
 -1814.1 & 1 & \makecell{\texttt{\_:ub1bL474C18 owl:onProperty ?n1.} } \\ 
 -1816.0 & 1 & \makecell{\texttt{\_:ub1bL422C18 rdf:type ?n1.} } \\ 
 -1802.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id28instance ?p1 swrs:ResearchTopic.} } \\ 
 -1780.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type swrs:ResearchTopic.} } \\ 
 -1811.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance rdf:type ?n1.} } \\ 
 -1799.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id54instance ?p1 swrs:ResearchTopic.} } \\ 
 -1804.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance ?p1 swrs:Person.} } \\ 
 -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id583instance rdf:type swrs:Person.} } \\ 
 -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance rdf:type swrs:Person.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2151instance ?p1 swrs:Person.} } \\ 
 -1796.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1116instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1777.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance swrs:isAbout http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1796.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id1073instance ?p1 http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id52instance.} } \\ 
 -1822.2 & 1 & \makecell{\texttt{swrs:name ?p2 ?n1.} } \\ 
 -1787.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance rdf:type swrs:InProceedings.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id663instance ?p1 swrs:InProceedings.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id709instance rdf:type swrs:Person.} } \\ 
 -1784.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance rdf:type swrs:Person.} } \\ 
 -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id655instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id848instance ?p1 swrs:Person.} } \\ 
 -1816.0 & 1 & \makecell{\texttt{\_:ub1bL355C18 rdf:type ?n1.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance ?p1 swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type swrs:Person.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id364instance rdf:type ?n1.} } \\ 
 -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id810instance ?p1 swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance ?p1 swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type swrs:Person.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id179instance rdf:type ?n1.} } \\ 
 -1805.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance ?p1 swrs:InProceedings.} } \\ 
 -1785.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id594instance rdf:type swrs:InProceedings.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id820instance ?p1 swrs:Person.} } \\ 
 -1807.3 & 1 & \makecell{\texttt{\_:ub1bL560C18 ?p1 swrs:Person.} } \\ 
 -1804.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id184instance ?p1 swrs:Person.} } \\ 
 -1786.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance rdf:type swrs:InProceedings.} } \\ 
 -1805.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id841instance ?p1 swrs:InProceedings.} } \\ 
 -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id684instance rdf:type swrs:InProceedings.} } \\ 
 -1784.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Article.} } \\ 
 -1787.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance rdf:type swrs:Publication.} } \\ 
 -1806.2 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id333instance ?p1 swrs:Publication.} } \\ 
 -1806.7 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id954instance ?p1 swrs:Publication.} } \\ 
 -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id773instance ?p1 swrs:Publication.} } \\ 
 -1815.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id20instance rdf:type ?n1.} } \\ 
 -1803.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id350instance ?p1 swrs:Person.} } \\ 
 -1803.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance ?p1 swrs:Person.} } \\ 
 -1784.9 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id262instance rdf:type swrs:Person.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance ?p1 swrs:Person.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type ?n1.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id598instance rdf:type swrs:Person.} } \\ 
 -1813.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id68instance swrs:phone ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id305instance rdf:type ?n1.} } \\ 
 -1822.2 & 1 & \makecell{\texttt{swrs:title ?p2 ?n1.} } \\ 
 -1791.2 & 1 & \makecell{\texttt{swrs:title ?p1 owl:DatatypeProperty.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance rdf:type ?n1.} } \\ 
 -1800.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id877instance.} } \\ 
 -1812.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name ?n1.} } \\ 
 -1781.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1800.1 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id576instance ?p1 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1831.1 & 1 & \makecell{\texttt{?n1 ?p2 "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1812.1 & 1 & \makecell{\texttt{?n1 swrs:name "Sam Chapman"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1813.8 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Forschungsgebiete/viewForschungsgebietOWL/id79instance swrs:name ?n1.} } \\ 
 -1811.1 & 1 & \makecell{\texttt{?n1 swrs:name "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1830.1 & 1 & \makecell{\texttt{?n1 ?p2 "CKDD - Conceptual Knowledge Discovery"^^<http://www.w3.org/2001/XMLSchema\_string>.} } \\ 
 -1808.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id165instance swrs:name ?n1.} } \\ 
 -1785.0 & 1 & \makecell{\texttt{\_:ub1bL199C18 rdf:type owl:Restriction.} } \\ 
 -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id18instance rdf:type ?n1.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance rdf:type swrs:Publication.} } \\ 
 -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id214instance ?p1 swrs:Publication.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id593instance rdf:type ?n1.} } \\ 
 -1786.6 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id499instance rdf:type swrs:InProceedings.} } \\ 
 -1805.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id275instance ?p1 swrs:InProceedings.} } \\ 
 -1821.4 & 1 & \makecell{\texttt{swrs:member ?p2 ?n1.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2118instance rdf:type swrs:Person.} } \\ 
 -1814.8 & 1 & \makecell{\texttt{swrs:Organization rdf:type ?n1.} } \\ 
 -1815.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id814instance rdf:type ?n1.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id439instance rdf:type ?n1.} } \\ 
 -1785.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id637instance rdf:type swrs:Person.} } \\ 
 -1788.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id703instance rdf:type swrs:Person.} } \\ 
 -1787.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance rdf:type swrs:Person.} } \\ 
 -1806.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id2008instance ?p1 swrs:Person.} } \\ 
 -1819.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id804instance rdf:type ?n1.} } \\ 
 -1803.3 & 1 & \makecell{\texttt{swrs:Proceedings ?p1 owl:Class.} } \\ 
 -1807.0 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewExternerAutorOWL/id786instance ?p1 swrs:Person.} } \\ 
 -1806.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance ?p1 swrs:Person.} } \\ 
 -1787.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type swrs:Person.} } \\ 
 -1818.5 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Personen/viewPersonOWL/id48instance rdf:type ?n1.} } \\ 
 -1787.3 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance rdf:type swrs:Publication.} } \\ 
 -1806.4 & 1 & \makecell{\texttt{http://www.aifb.uni-karlsruhe.de/Publikationen/viewPublikationOWL/id542instance ?p1 swrs:Publication.} } \\ 
\hline
\end{longtable}

\subsection{Mutag, top 100 by log factor}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 11178.4 & 5996 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom ?n2.} } \\ 
 1014.6 & 1840 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 938.0 & 1850 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 577.6 & 1897 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 -1250.9 & 138 & \makecell{\texttt{?n1 rdfs:subClassOf ?n2.} } \\ 
 -1422.9 & 320 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Ar\_halide.} } \\ 
 -1593.5 & 319 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_charge ?n2.} } \\ 
 -1689.4 & 181 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -1707.3 & 173 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1719.3 & 181 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -1737.1 & 173 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1830.8 & 110 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -1859.3 & 110 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -1865.4 & 142 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
 -1879.4 & 22 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -1886.3 & 19 & \makecell{\texttt{?n1 rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -1894.6 & 142 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
 -1911.7 & 67 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -1928.9 & 50 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -1938.7 & 67 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -1981.2 & 36 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d219 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -1988.7 & 16 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d84 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1996.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond5500 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d270\_13.} } \\ 
 -2000.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -2000.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2001.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_6.} } \\ 
 -2002.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond100 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-318 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2009.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Molybdenum rdf:type owl:Class.} } \\ 
 -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 rdf:type owl:Class.} } \\ 
 -2011.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond32 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d10\_6.} } \\ 
 -2012.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3442 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond16 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2014.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -2017.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p1 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -2018.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p1 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2018.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p1 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
 -2019.5 & 25 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d216 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p1 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3218 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2020.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2022.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2025.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-120 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-980 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-537 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1155 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-946 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ar\_halide ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Alcohol ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 ?p1 owl:Class.} } \\ 
 -2030.2 & 4 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2031.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -2031.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 rdfs:subClassOf ?n1.} } \\ 
 -2032.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p1 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -2032.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2033.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf ?n1.} } \\ 
 -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Iodine-95 rdfs:subClassOf ?n1.} } \\ 
 -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 rdfs:subClassOf ?n1.} } \\ 
 -2034.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2041.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
 -2042.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2042.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2491 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-222 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-1445 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 rdf:type ?n1.} } \\ 
 -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 rdf:type ?n1.} } \\ 
 -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Nitrogen-499 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-1525 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-2912 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-3038 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-216 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 rdf:type ?n1.} } \\ 
 -2044.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 http://dl-learner.org/carcinogenesis\_charge ?n1.} } \\ 
 -2045.1 & 1 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2045.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d82\_26 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1421 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4108 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond7147 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond566 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3448 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond662 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5306 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5197 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond6533 rdf:type ?n1.} } \\ 
 -2048.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type ?n1.} } \\ 
 -2052.2 & 6 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2053.6 & 66 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3169 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2056.6 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p2 ?n1.} } \\ 
 -2059.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p2 ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
 -2059.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p2 ?n1.} } \\ 
 -2059.9 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring ?p2 ?n1.} } \\ 
 -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
 -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 ?p2 ?n1.} } \\ 
 -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p2 ?n1.} } \\ 
 -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p2 ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p2 ?n1.} } \\ 
 -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di51 ?p2 ?n1.} } \\ 
 -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 ?p2 ?n1.} } \\ 
 -2061.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 ?p2 ?n1.} } \\ 
 -2062.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p2 ?n1.} } \\ 
 -2063.1 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 ?p2 ?n1.} } \\ 
 -2065.9 & 9 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d104 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -2071.1 & 2 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
 -2071.4 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2071.9 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2073.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p2 ?n1.} } \\ 
 -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p2 ?n1.} } \\ 
 -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p2 ?n1.} } \\ 
 -2077.4 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 ?p2 ?n1.} } \\ 
 -2086.2 & 3 & \makecell{\texttt{?n1 rdf:type ?n2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d11\_4.} } \\ 
 -2094.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
 -2094.6 & 541 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -2114.7 & 92 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond ?n2.} } \\ 
 -2139.2 & 4 & \makecell{\texttt{?n1 ?p3 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -2151.0 & 62 & \makecell{\texttt{?n1 ?p4 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -2743.8 & 126 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
\hline
\end{longtable}

\subsection{Mutag, top 100 by frequency}

\begin{longtable}{ r r p{19cm} }
 log factor & frequency & \\
\hline\endhead
 11178.4 & 5996 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom ?n2.} } \\ 
 577.6 & 1897 & \makecell{\texttt{?n1 rdf:type ?n2.} } \\ 
 938.0 & 1850 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 1014.6 & 1840 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2094.6 & 541 & \makecell{\texttt{?n1 ?p3 ?n2.} } \\ 
 -1422.9 & 320 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Ar\_halide.} } \\ 
 -1593.5 & 319 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_charge ?n2.} } \\ 
 -1719.3 & 181 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -1689.4 & 181 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -1707.3 & 173 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1737.1 & 173 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -1894.6 & 142 & \makecell{\texttt{?n1 ?p2 owl:Class.} } \\ 
 -1865.4 & 142 & \makecell{\texttt{?n1 rdf:type owl:Class.} } \\ 
 -1250.9 & 138 & \makecell{\texttt{?n1 rdfs:subClassOf ?n2.} } \\ 
 -2743.8 & 126 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -1859.3 & 110 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -1830.8 & 110 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2114.7 & 92 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond ?n2.} } \\ 
 -1911.7 & 67 & \makecell{\texttt{?n1 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -1938.7 & 67 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2053.6 & 66 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3169 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2151.0 & 62 & \makecell{\texttt{?n1 ?p4 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 ?p3 ?n2.} } \\ 
 -1928.9 & 50 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -1981.2 & 36 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d219 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -2019.5 & 25 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d216 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -1879.4 & 22 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -1886.3 & 19 & \makecell{\texttt{?n1 rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -1988.7 & 16 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d84 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2065.9 & 9 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d104 http://dl-learner.org/carcinogenesis\_hasStructure ?n1.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -2052.2 & 6 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2071.4 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2071.9 & 5 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2030.2 & 4 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2139.2 & 4 & \makecell{\texttt{?n1 ?p3 http://dl-learner.org/carcinogenesis\_d11\_4.} \\\texttt{?n1 rdf:type ?n2.} } \\ 
 -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p2 ?n1.} } \\ 
 -2032.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2073.9 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p2 ?n1.} } \\ 
 -2086.2 & 3 & \makecell{\texttt{?n1 rdf:type ?n2.} \\\texttt{?n1 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d11\_4.} } \\ 
 -2075.7 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p2 ?n1.} } \\ 
 -2079.2 & 3 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 ?p2 ?n1.} } \\ 
 -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2036.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2031.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2035.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2071.1 & 2 & \makecell{\texttt{?n1 ?p2 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
 -2062.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p2 ?n1.} } \\ 
 -2063.1 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 ?p2 ?n1.} } \\ 
 -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di51 ?p2 ?n1.} } \\ 
 -2056.6 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p2 ?n1.} } \\ 
 -2059.7 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p2 ?n1.} } \\ 
 -2061.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 ?p2 ?n1.} } \\ 
 -2077.4 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 ?p2 ?n1.} } \\ 
 -2042.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 http://dl-learner.org/carcinogenesis\_inBond ?n1.} } \\ 
 -2094.0 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n2.} \\\texttt{?n2 ?p3 ?n1.} } \\ 
 -2059.9 & 2 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring ?p2 ?n1.} } \\ 
 -2045.1 & 1 & \makecell{\texttt{?n1 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 ?p1 http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond412 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d112\_4.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond100 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond95 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2001.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_6.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2012.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2026.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond99 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} \\\texttt{http://dl-learner.org/carcinogenesis\_d102 http://dl-learner.org/carcinogenesis\_hasAtom http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2003.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2020.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond101 ?p1 http://dl-learner.org/carcinogenesis\_d102\_2.} } \\ 
 -2018.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1012 ?p1 http://dl-learner.org/carcinogenesis\_d135\_23.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-980 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1585 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1421 rdf:type ?n1.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-120 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4108 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond7147 rdf:type ?n1.} } \\ 
 -2017.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester ?p1 http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -2000.2 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf http://dl-learner.org/carcinogenesis\_Structure.} } \\ 
 -2033.6 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ester rdfs:subClassOf ?n1.} } \\ 
 -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Iodine-95 rdfs:subClassOf ?n1.} } \\ 
 -2034.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Chlorine-93 rdfs:subClassOf ?n1.} } \\ 
 -2045.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d82\_26 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-1525 rdf:type ?n1.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2172 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Nitrogen-499 rdf:type ?n1.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-537 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p1 http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 ?p2 ?n1.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type http://dl-learner.org/carcinogenesis\_Alcohol.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-2134 rdf:type ?n1.} } \\ 
 -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 rdf:type ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p2 ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2759 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2009.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2182 rdf:type http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond566 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-2912 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2025.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2022.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2154 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2026.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2008.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2041.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 rdf:type ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
 -2059.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-3065 ?p2 ?n1.} \\\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
 -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p2 ?n1.} } \\ 
 -2026.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 ?p1 http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-320 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type ?n1.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} } \\ 
 -2020.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 ?p1 owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2002.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Phosphorus-60 rdf:type owl:Class.} \\\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2000.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -1996.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type http://dl-learner.org/carcinogenesis\_Phosphorus-60.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond5500 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d270\_13.} } \\ 
 -2018.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 ?p1 http://dl-learner.org/carcinogenesis\_Phosphorus-60.} } \\ 
 -2034.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d270\_13 rdf:type ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 ?p2 ?n1.} } \\ 
 -2042.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2634 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond6533 rdf:type ?n1.} } \\ 
 -2048.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type ?n1.} } \\ 
 -2014.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 rdf:type http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -2032.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond9012 ?p1 http://dl-learner.org/carcinogenesis\_Bond-1.} } \\ 
 -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Molybdenum rdf:type owl:Class.} } \\ 
 -2031.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Di260 rdfs:subClassOf ?n1.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond485 rdf:type ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p2 ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-939 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3448 rdf:type ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2491 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-3038 rdf:type ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1155 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p2 ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-1260 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-946 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Ar\_halide ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Alcohol ?p1 owl:Class.} } \\ 
 -2027.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 ?p1 owl:Class.} } \\ 
 -2010.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Carbon-22 rdf:type owl:Class.} } \\ 
 -2044.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_d243\_16 http://dl-learner.org/carcinogenesis\_charge ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-216 rdf:type ?n1.} } \\ 
 -2059.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 ?p2 ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-2916 rdf:type ?n1.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond1962 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond3442 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2030.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond4893 ?p1 http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond2380 rdf:type ?n1.} } \\ 
 -2009.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-318 rdf:type http://dl-learner.org/carcinogenesis\_Alkyl\_halide.} } \\ 
 -2042.1 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_di10-3291 rdf:type ?n1.} } \\ 
 -2043.9 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 rdf:type ?n1.} } \\ 
 -2061.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_ar\_halide-180 ?p2 ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-222 rdf:type ?n1.} } \\ 
 -2043.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_Non\_ar\_6c\_ring rdf:type ?n1.} } \\ 
 -2060.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 ?p2 ?n1.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alcohol-1069 rdf:type ?n1.} } \\ 
 -2060.3 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p2 ?n1.} } \\ 
 -2026.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amine-2275 ?p1 http://dl-learner.org/carcinogenesis\_Amine.} } \\ 
 -2042.5 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_alkyl\_halide-1445 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type ?n1.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2011.7 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond32 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} \\\texttt{http://dl-learner.org/carcinogenesis\_bond32 http://dl-learner.org/carcinogenesis\_inBond http://dl-learner.org/carcinogenesis\_d10\_6.} } \\ 
 -2013.0 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond16 rdf:type http://dl-learner.org/carcinogenesis\_Bond-7.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond662 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5306 rdf:type ?n1.} } \\ 
 -2046.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_bond5197 rdf:type ?n1.} } \\ 
 -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3218 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2020.8 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 ?p1 http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
 -2003.4 & 1 & \makecell{\texttt{http://dl-learner.org/carcinogenesis\_amino-3151 rdf:type http://dl-learner.org/carcinogenesis\_Amino.} } \\ 
\hline
\end{longtable}

\end{landscape}
\end{document}